# Tables

## Variables

(ref:cap-variables) **Set of explicative variables used to model the spatial probability of deforestation**. A total of height variables were tested. They described topography, forest accessibility, forest landscape, land tenure and deforestation history.\vspace{0.5cm} 

```{r variables, results="asis", echo=FALSE}
df.variables <- read.table("tables/variables.txt", header=TRUE, sep=",")
vect.align <- c(rep("l",3), rep("r",2))
col.names <- c("Product","Source","Variable derived","Unit","Resolution (m)")
knitr::kable(df.variables, caption="(ref:cap-variables)", booktabs=TRUE,
			 col.names=col.names,
			 align=vect.align, linesep="") %>%
	column_spec(c(1:3),width="3cm") %>%
	column_spec(4,width="2cm") %>%
	column_spec(5,width="3cm") %>%
	kable_styling(latex_options=c("HOLD_position","striped"), full_width=FALSE, position="center")
```

## Mathematical formulas for accuracy indices

(ref:cap-confusion-matrix) **Confusion matrix used to compute accuracy indices**. A confusion matrix can be computed to compare model predictions with observations.\vspace{0.5cm} 

```{r confusion-matrix, results="asis", echo=FALSE}
df <- read.table("tables/confusion-matrix.txt", header=TRUE, sep=",")
vect.align <- rep("l",2)
col.names <- c("","","Observations","","Total")
options(knitr.kable.NA="")
knitr::kable(df, caption="(ref:cap-confusion-matrix)", booktabs=TRUE,
			 escape=FALSE,
			 col.names=col.names,
			 align=vect.align, linesep="") %>%
	kable_styling(latex_options=c("HOLD_position","striped"),
				  full_width=FALSE, position="center")
```


(ref:cap-accuracy-indices) **Formulas used to compute accuracy indices**. Several accuracy indices can be computed from the confusion matrix to estimate and compare models' predictive skill. We followed the definitions of @Pontius2008 for the FOM and @Liu2011 for the other indices. Note that the AUC relies on the predicted probabilities for observations 0 (non-deforested) and 1 (deforested), not on the confusion matrix. \vspace{0.5cm} 

```{r accuracy-indices, results="asis", echo=FALSE}
df <- read.table("tables/accuracy-indices.txt", header=TRUE, sep=",")
vect.align <- rep("l",2)
col.names <- c("Index","Formula")
knitr::kable(df, caption="(ref:cap-accuracy-indices)", booktabs=TRUE,
			 escape=FALSE,
			 col.names=col.names,
			 align=vect.align, linesep="") %>%
	kable_styling(latex_options=c("HOLD_position","striped"), full_width=FALSE, position="center")
```

\newpage
