# Tables

```{r lib-tab, echo=FALSE, results="hide"}
library(dplyr)
```

## Categories of the forest cover annual product

(ref:cap-cat-annual-product) **Categories of the forest cover annual product by @Vancutsem2020**. The forest cover annual product classifies Landsat image pixels in 16 categories for each year (on the 31$^{\text{st}}$ of December) between 1982 and 2019 and allows identifying moist tropical forest pixels at each date. Water data comes from @Pekel2016.\vspace{0.5cm} 

```{r cat-annual-product, results="asis", echo=FALSE}
df.cat <- read.table("tables/cat-annual-product.txt", header=TRUE, sep=";")
vect.align <- c("l", "r")
col.names <- c("Class", "Definition")
knitr::kable(df.cat, caption="(ref:cap-cat-annual-product)", booktabs=TRUE,
						 escape=FALSE,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
	column_spec(1,width="2cm") %>%
	column_spec(2,width="10cm") %>%
	kable_styling(latex_options=c("HOLD_position","striped"), full_width=FALSE, position="center")
```

## Variables

(ref:cap-variables) **Set of explicative variables used to model the spatial probability of deforestation**. A total of height variables were tested. They indicate topography, forest accessibility, forest landscape, deforestation history, and conservation status.\vspace{0.5cm} 

```{r variables, results="asis", echo=FALSE}
df.variables <- read.table("tables/variables.txt", header=TRUE, sep=",")
vect.align <- c(rep("l",3), rep("r",3))
col.names <- c("Product","Source","Variable derived","Unit","Resolution (m)","Date")
knitr::kable(df.variables, caption="(ref:cap-variables)", booktabs=TRUE,
						 escape=FALSE,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
	column_spec(c(1:3),width="3cm") %>%
	column_spec(4,width="2cm") %>%
	column_spec(5,width="3cm") %>%
	column_spec(5,width="2cm") %>%
	kable_styling(latex_options=c("HOLD_position","striped"), full_width=FALSE, position="center")
```

## Sample size

(ref:cap-samp-size) **Number of observations used for the spatial model of deforestation for each study area**. The table includes the number of non-deforested (nfor) and deforested (ndef) pixels per study area. These numbers include the forest pixels with full information regarding the explanatory variables. The corresponding number of hectares is also provided (nfHa and ndHa, respectively).\vspace{0.5cm} 

```{r samp-size, results="asis", echo=FALSE}
df.cat <- read.table("tables/samp_size.csv", header=TRUE, sep=",")
df.cat <- df.cat[,-c(1)]
vect.align <- c(rep("l", 2), rep("r", 4))
col.names <- c("Country -- Study-area","Code","nfor","ndef","nfHa","ndHa")
knitr::kable(df.cat, caption="(ref:cap-samp-size)", booktabs=TRUE, longtable=TRUE,
						 format.args=list(big.mark=","),
						 escape=FALSE,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  pack_rows("All continents", 120, 120) %>%
	kable_styling(latex_options=c("HOLD_position","striped","repeat_header"), full_width=FALSE,
	              position="center", font_size=11, repeat_header_method="replace")
```

## Mathematical formulas for accuracy indices

(ref:cap-confusion-matrix) **Confusion matrix used to compute accuracy indices**. A confusion matrix can be computed to compare model predictions with observations.\vspace{0.5cm} 

```{r confusion-matrix, results="asis", echo=FALSE}
df <- read.table("tables/confusion-matrix.txt", header=TRUE, sep=",")
vect.align <- rep("l",2)
col.names <- c("","","Observations","","Total")
options(knitr.kable.NA="")
knitr::kable(df, caption="(ref:cap-confusion-matrix)", booktabs=TRUE,
						 escape=FALSE,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
	kable_styling(latex_options=c("HOLD_position","striped"),
				  full_width=FALSE, position="center")
```


(ref:cap-accuracy-indices) **Formulas used to compute accuracy indices**. Several accuracy indices can be computed from the confusion matrix to estimate and compare models' predictive skill. We followed the definitions of @Pontius2008 for the FOM and @Liu2011 for the other indices. Note that the AUC relies on the predicted probabilities for observations 0 (non-deforested) and 1 (deforested), not on the confusion matrix.\vspace{0.5cm} 

```{r accuracy-indices, results="asis", echo=FALSE}
df <- read.table("tables/accuracy-indices.txt", header=TRUE, sep=",")
vect.align <- rep("l",2)
col.names <- c("Index","Formula")
knitr::kable(df, caption="(ref:cap-accuracy-indices)", booktabs=TRUE,
						 escape=FALSE,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
	kable_styling(latex_options=c("HOLD_position","striped"), full_width=FALSE, position="center")
```

## Accuracy indices

(ref:cap-accuracy) **Accuracy indices' weighted averages for the three statistical models**. Accuracy indices were averaged accross study areas using forest cover areas in 2010 as weights. D: percentage of deviance explained, AUC Area Under ROC Curve, OA: overall accuracy, FOM: Figure Of Merite, TSS: True Skill Statistics. Averaged accuracy indices were computed for the three statistical models: "glm", "icar", and "rf" model.\vspace{0.5cm} 

```{r accuracy, results="asis", echo=FALSE}
df <- read.table("tables/perf_mod.csv", header=TRUE, sep=",")
vect.align <- c(rep("l", 1), rep("r", 5))
col.names <- c("Model","D","AUC","OA","FOM","TSS")
knitr::kable(df, caption="(ref:cap-accuracy)", booktabs=TRUE,
						 escape=FALSE,
						 digits=1,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
	kable_styling(latex_options=c("HOLD_position","striped"), stripe_index=c(2), full_width=FALSE, position="center")
```


(ref:cap-accuracy-cont) **Accuracy indices' weighted averages for the three statistical models by continent**. Accuracy indices were averaged accross study areas for each continent using forest cover areas in 2010 as weights. D: percentage of deviance explained, AUC Area Under ROC Curve, OA: overall accuracy, FOM: Figure Of Merite, TSS: True Skill Statistics. Averaged accuracy indices were computed for the three statistical models: "glm", "icar", and "rf" model.\vspace{0.5cm} 

```{r accuracy-cont, results="asis", echo=FALSE}
df <- read.table("tables/perf_cont_mod.csv", header=TRUE, sep=",")
vect.align <- c(rep("l", 1), rep("r", 5))
col.names <- c("Continent","Model","D","AUC","OA","FOM","TSS")
knitr::kable(df, caption="(ref:cap-accuracy-cont)", booktabs=TRUE,
						 escape=FALSE,
						 digits=1,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
  column_spec(1, bold=TRUE) %>%
  collapse_rows(columns=1, latex_hline = "major", valign = "middle") %>%
  kable_styling(latex_options=c("HOLD_position","striped"), stripe_index=c(2,5,8), full_width=FALSE, position="center")
```

## Historical forest cover change

(ref:cap-fcc-hist) **Historical forest cover change for each study-area**. Forest cover areas are given in thousand hectares (Kha) for the years 2000, 2010 and 2020 ("fc2000", "fc2010", and "fc2020", respectively). The mean annual deforested area $d$ for the period 2010--2020 is given in hectare per year (ha/yr). For comparing the deforestation intensity between study-areas, the corresponding mean annual deforestation rate $p$ on the period 2010--2020 is also provided in percent per year (%/yr), with one decimal precision.\vspace{0.5cm} 

```{r fcc-hist, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/forest_cover_change.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::select(area_name, for2000, for2010, for2020, andef, pdef) %>%
  mutate_at(.var=vars(starts_with("for")),.fun=function(x){round(x/1000)})
## Make table
vect.align <- c(rep("l", 1), rep("r", 5))
col.names <- c("Country -- Study-area", "fc2000", "fc2010", "fc2020", "$d$ (ha/yr)", "$p$ (\\%/yr)")
knitr::kable(df, caption="(ref:cap-fcc-hist)", booktabs=TRUE, longtable=TRUE,
						 format.args=list(big.mark=","),
						 escape=FALSE,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
	kable_styling(latex_options=c("HOLD_position","striped","repeat_header"), full_width=FALSE,
	              position="center", font_size=11, repeat_header_method="replace")
```

(ref:cap-fcc-hist-reg) **Historical forest cover change per region and continent**. Areas of forest cover are given in thousand hectares (Kha) for the years 2000, 2010 and 2020 ("fc2000", "fc2010", and "fc2020", respectively). The mean annual deforested area $d$ for the period 2010--2020 is given in hectare per year (ha/yr). For comparing the deforestation intensity between study-areas, the corresponding mean annual deforestation rate $p$ on the period 2010--2020 is also provided in percent per year (%/yr), with one decimal precision. Estimates for America include Brazil, and estimates for Asia include India. Around 7.5 Mha (75,000 km$^2$, about the size of Scotland or South Carolina) of natural old-growth moist tropical forest have been disappearing each year in the period 2010--2020.\vspace{0.5cm} 

```{r fcc-hist-reg, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/fcc_hist_region.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>% 
  dplyr::select(area_cont, for2000, for2010, for2020, andef, pdef) %>%
  mutate_at(.var=vars(starts_with("for")),.fun=function(x){round(x/1000)})
## Make table
vect.align <- c(rep("l", 1), rep("r", 5))
col.names <- c("Region", "fc2000", "fc2010", "fc2020", "$d$ (ha/yr)", "$p$ (\\%/yr)")
knitr::kable(df, caption="(ref:cap-fcc-hist-reg)", booktabs=TRUE, longtable=FALSE,
						 format.args=list(big.mark=","),
						 escape=FALSE,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
	kable_styling(latex_options=c("HOLD_position","striped","repeat_header"), full_width=FALSE,
	              position="center", font_size=11, repeat_header_method="replace")
```

\newpage

## Forest cover projections

(ref:cap-fcc-proj) **Forest cover projections for each study-area**. Projected areas of forest cover are given in thousand hectares (Kha) for four years in the future (2040, 2060, 2080, and 2100). Projections were made using the forest cover in 2020 and the mean annual deforested area on the period 2010--2020 ("fc2000" and $d$ respectively in Table \@ref(tab:fcc-hist)), assuming a "business-as-usual" scenario of deforestation. Column "loss21" indicates the projected percentage of forest cover loss during the 21$^\text{st}$ century (2100 vs. 2000). Column "yrdis" indicates the estimated year at which all the forest of the study-area will have disappeared.\vspace{0.5cm} 

```{r fcc-proj, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/forest_cover_change.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::mutate(yrdis=as.character(yrdis)) %>%
  dplyr::select(area_name, for2040, for2060, for2080, for2100, loss21, yrdis) %>%
  mutate_at(.var=vars(starts_with("for")),.fun=function(x){round(x/1000)})
## Make table
vect.align <- c(rep("l", 1), rep("r", 6))
col.names <- c("Country -- Study-area", "fc2040", "fc2060", "fc2080", "fc2100", "loss21 (\\%)", "yrdis")
knitr::kable(df, caption="(ref:cap-fcc-proj)", booktabs=TRUE, longtable=TRUE,
						 format.args=list(big.mark=","),
						 escape=FALSE,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
	kable_styling(latex_options=c("HOLD_position","striped","repeat_header"), full_width=FALSE,
	              position="center", font_size=11, repeat_header_method="replace")
```

(ref:cap-fcc-proj-reg) **Forest cover projections per region and continent**. Projected areas of forest cover are given in thousand hectares (Kha) for four dates in the future (2040, 2060, 2080, and 2100). Projections were made using the forest cover in 2020 and the mean annual deforested area on the period 2010--2020 ("fc2000" and $d$ respectively in Table \@ref(tab:fcc-hist-reg)), assuming a "business-as-usual" scenario of deforestation. Column "loss21" indicates the projected percentage of forest cover loss during the 21$^\text{st}$ century (2100 vs. 2000). At the continental level, it makes less sense to compute the year at which all the forest will have disappeared, as some countries might conserve forest for a very long time, even though they account for a very small proportion of the total forest area at the continental scale. Instead, we computed the estimated year at which 75% of the forest cover in 2000 will have disappeared ("yr75dis").\vspace{0.5cm} 

```{r fcc-proj-reg, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/fcc_proj_region.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::mutate(yr75dis=as.character(yr75dis)) %>%
  dplyr::select(cont, for2040, for2060, for2080, for2100, loss21, yr75dis) %>%
  mutate_at(.var=vars(starts_with("for")),.fun=function(x){round(x/1000)})
## Make table
vect.align <- c(rep("l", 1), rep("r", 6))
col.names <- c("Country -- Study-area", "fc2040", "fc2060", "fc2080", "fc2100", "loss21 (\\%)", "yr75dis")
knitr::kable(df, caption="(ref:cap-fcc-proj-reg)", booktabs=TRUE, longtable=FALSE,
						 format.args=list(big.mark=","),
						 digits=0,
						 escape=FALSE,
						 col.names=col.names,
						 align=vect.align, linesep="") %>%
	kable_styling(latex_options=c("HOLD_position","striped","repeat_header"), full_width=FALSE,
	              position="center", font_size=11, repeat_header_method="replace")
```

\newpage
