# Supplementary tables

```{r lib-tab, echo=FALSE, results="hide"}
library(dplyr)
```

<!--------------------------------------------------->
<!-- Categories of the forest cover annual product -->
<!--------------------------------------------------->

## Table S1 -- Pixel categories of the forest cover annual product {-}

(ref:cap-cat-annual-product) **Pixel categories of the forest cover annual product by @Vancutsem2021**. The forest cover annual product v1_2020 classifies Landsat image pixels in 6 categories for each year (on the 1$^{\text{st}}$ of January) between 1990 and 2021 and allows identifying tropical moist forest pixels at each date.\vspace{0.5cm}

```{r cat-annual-product, results="asis", echo=FALSE}
df.cat <- read.table("tables/cat-annual-product.txt", header=TRUE, sep=";")
vect.align <- c("l", "r")
col.names <- c("Class", "Definition")
knitr::kable(df.cat, caption="(ref:cap-cat-annual-product)", booktabs=TRUE,
                         escape=FALSE,
                         col.names=col.names,
                         align=vect.align, linesep="") %>%
    column_spec(1,width="2cm") %>%
    column_spec(2,width="10cm") %>%
    kable_styling(latex_options=c("HOLD_position","striped"), full_width=FALSE, position="center")
```

<!--------------->
<!-- Variables -->
<!--------------->

## Table S2 -- Variables {-}

(ref:cap-variables) **Set of explanatory variables used to model the spatial probability of deforestation**. A total of height variables were tested. Variables describe topography, forest accessibility, forest landscape, deforestation history, and conservation status.\vspace{0.5cm}

```{r variables, results="asis", echo=FALSE}
df.variables <- read.table("tables/variables.txt", header=TRUE, sep=",")
vect.align <- c(rep("l",3), rep("r",3))
col.names <- c("Product","Source","Version / Date","Variable derived","Unit","Resolution (m)")
knitr::kable(df.variables, caption="(ref:cap-variables)", booktabs=TRUE,
                         escape=FALSE,
                         col.names=col.names,
                         align=vect.align, linesep="") %>%
    column_spec(c(1:2),width="2.5cm") %>%
    column_spec(3,width="2cm") %>%
    column_spec(4,width="2.5cm") %>%
    column_spec(c(5:6),width="1.5cm") %>%
    kable_styling(latex_options=c("HOLD_position","striped"), full_width=FALSE, position="center", font_size=11)
```

\newpage

<!----------------->
<!-- Sample size -->
<!----------------->

## Table S3 -- Sample size {-}

(ref:cap-samp-size) **Number of observations used for the spatial model of deforestation for each study area**. The table includes the number of non-deforested ("nfor") and deforested ("ndef") pixels per study area. These numbers include the forest pixels with full information regarding the explanatory variables. The corresponding number of hectares is also provided ("nfHa" and "ndHa", respectively).\vspace{0.5cm}

```{r samp-size, results="asis", echo=FALSE}
df.cat <- read.table("tables/samp_size.csv", header=TRUE, sep=",")
df.cat <- df.cat[,-c(1)]
vect.align <- c(rep("l", 2), rep("r", 4))
col.names <- c("Country -- Study area","Code","nfor","ndef","nfHa","ndHa")
unit.vect <- c("", "", "", "", "(ha)", "(ha)")
knitr::kable(df.cat, caption="(ref:cap-samp-size)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=FALSE, align=vect.align) %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  pack_rows("All continents", 120, 120) %>%
    kable_styling(latex_options=c("HOLD_position","striped","repeat_header"), full_width=FALSE,
                  position="center", font_size=11, repeat_header_method="replace")
```

\newpage

<!------------------------------------------------->
<!-- Parameter estimates and variable importance -->
<!------------------------------------------------->

## Tables S4--S5 -- Parameter estimates and variable importance {-}

(ref:cap-par) **Parameter estimates for each study area**. For each study area, we computed the posterior mean of each parameter ("int": intercept, "pa": protected area effect, "elev", "slope", "ddefor","dedge", "driver", "droad", "dtown": slope parameters associated to elevation, slope, distance to past deforestation, distance to forest edge, distance to nearest river, distance to nearest road, and distance to nearest town, respectively, "Vrho": variance of the spatial random effects). Continuous explanatory variables were normalized (mean=0 and standard-deviation=1), allowing us to estimate the relative importance of each variable in determining the spatial probability of deforestation from parameter values. Comparison can be done within and between study areas.\vspace{0.5cm}

```{r par, results="asis", echo=FALSE}
options(knitr.kable.NA="--")
## Import data
df <- read.table("tables/parameter_estimates.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::select(-area_cont, -area_ctry, -area_name)
## Make table
vect.align <- c(rep("l", 1), rep("r", 10))
col.names <- c("study area", "int", "pa", "elev", "slope",
               "ddefor", "dedge", "driver", "droad", "dtown", "Vrho")
knitr::kable(df, caption="(ref:cap-par)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         digits=3,
                         escape=FALSE,
                         col.names=col.names,
                         align=vect.align, linesep="") %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=10, repeat_header_method="replace")
```

\newpage

(ref:cap-par-region) **Parameter estimate weighted means per region**. We used the forest cover in 2010 to compute the parameter estimate weighted mean per region. Continuous explanatory variables were normalized (mean=0 and standard-deviation=1), allowing us to estimate the relative importance of each variable in determining the spatial probability of deforestation from parameter values. Comparison can be done within and between regions. At pantropical scale (when considering all continents together), explanatory variables can be classified in the following decreasing order of importance: distance to forest edge, distance to past deforestation, presence of a protected area, distance to nearest road, elevation, distance to nearest town, slope, and distance to nearest river.\vspace{0.5cm}

```{r par-region, results="asis", echo=FALSE}
options(knitr.kable.NA="--")
## Import data
df <- read.table("tables/weighted_param_region.csv", header=TRUE, sep=",")
## Make table
vect.align <- c(rep("l", 1), rep("r", 10))
col.names <- c("Region", "int", "pa", "elev", "slope",
               "ddefor", "dedge", "driver", "droad", "dtown", "Vrho")
knitr::kable(df, caption="(ref:cap-par-region)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         digits=3,
                         escape=FALSE,
                         col.names=col.names,
                         align=vect.align, linesep="") %>%
    kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                  full_width=FALSE,
                  position="center", font_size=10, repeat_header_method="replace")
```

\newpage

<!------------------------------------------------>
<!-- Effect of protected areas on deforestation -->
<!------------------------------------------------>

## Table S6 -- Effect of protected areas on deforestation {-}

```{r pa-sign, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/sign_PA_road.csv", header=TRUE, sep=",")
nsa <- df$nctry[df$var=="PA"]
nsignif <- df$nctry_sign[df$var=="PA"]
perc <- df$perc[df$var=="PA"]
perc_w <- df$perc_w[df$var=="PA"]
```

(ref:cap-pa) **Effect of protected areas on deforestation**. We show here the estimated effect of the presence of a protected area on the probability of deforestation for each study area. We computed the mean ("Mean"), the standard-deviation ("Sd"), and the bayesian 95% credible interval ("CI 95%") of the estimated parameter. Column "P" indicates the percentage of decrease in the deforestation risk inside protected areas. Column "signif" indicates (with a star) that the estimated effect was negative and significantly different from zero (zero not included in the credible interval). Out of the `r nsa` study areas, `r nsignif` showed a significant negative effect (`r perc`% of the countries). These `r nsignif` study areas accounted for `r perc_w`% of the moist tropical forest in 2010 ("fc2010" in Kha).\vspace{0.5cm}

```{r pa, results="asis", echo=FALSE}
options(knitr.kable.NA="--")
## Format function
f_format <- function(x) {
  if (is.na(x)) {
    r <- NA
  } else if (x >=0) {
    r <- format(paste0(" ", format(x, digits=1, nsmall=3)), width=6, justify="right")
  } else if (x <0) {
    r <- format(format(x, digits=1, nsmall=3), width=6, justify="right")
  }
  return(r)
}
## Import data
df <- read.table("tables/parea_estimates.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::mutate(CI=paste0("(",
                          sapply(df$CI_low, f_format),
                          ", ",
                          sapply(df$CI_high, f_format),
                          ")")) %>%
  dplyr::mutate(for2010=round(for2010/1000)) %>%
  dplyr::mutate(sign=ifelse(CI_low * CI_high > 0 & !is.na(Mean), 1, 0)) %>%
  dplyr::select(-area_cont, -area_ctry, -area_code, -CI_low, -CI_high) %>%
  dplyr::mutate(CI=ifelse(CI=="(NA, NA)", NA, CI)) %>%
  dplyr::mutate(sign=ifelse(sign==1, "$\\star$", "")) %>%
  dplyr::relocate(CI, .before=P)
## Make table
vect.align <- c(rep("l", 1), rep("r", 5), "c")
col.names <- c("Country -- study area", "fc2010", "Mean", "Sd", "CI 95%",
               "P", "signif")
unit.vect <- c("", "(Kha)", "", "", "", "\\%", "")
knitr::kable(df, caption="(ref:cap-pa)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         digits=3,
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=TRUE, align=vect.align) %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=11, repeat_header_method="replace")
```

\newpage

<!----------------------------------------------------->
<!-- Effect of the distance to road on deforestation -->
<!----------------------------------------------------->

## Table S7 -- Effect of the distance to road on deforestation {-}

```{r road-sign, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/sign_PA_road.csv", header=TRUE, sep=",")
nsa <- df$nctry[df$var=="road"]
nsignif <- df$nctry_sign[df$var=="road"]
perc <- df$perc[df$var=="road"]
perc_w <- df$perc_w[df$var=="road"]
```

(ref:cap-road) **Effect of the distance to road on deforestation**. We show here the estimated effect of the distance to the nearest road on the probability of deforestation for each study area. We computed the mean ("Mean"), the standard-deviation ("Sd"), and the bayesian 95% credible interval ("CI 95%") of the estimated parameter. Column "signif" indicates (with a star) that the estimated effect was negative and significantly different from zero (zero not included in the credible interval). Out of the `r nsa` study areas, `r nsignif` showed a significant negative effect (`r perc`% of the countries). These `r nsignif` study areas accounted for `r perc_w`% of the moist tropical forest in 2010 ("fc2010" in Kha).\vspace{0.5cm}

```{r road, results="asis", echo=FALSE}
options(knitr.kable.NA="--")
## Import data
df <- read.table("tables/road_estimates.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::mutate(CI=paste0("(",
                          sapply(df$CI_low, f_format),
                          ", ",
                          sapply(df$CI_high, f_format),
                          ")")) %>%
  dplyr::mutate(for2010=round(for2010/1000)) %>%
  dplyr::mutate(sign=ifelse(CI_low * CI_high > 0 & !is.na(Mean), 1, 0)) %>%
  dplyr::select(-area_cont, -area_ctry, -area_code, -CI_low, -CI_high) %>%
  dplyr::mutate(CI=ifelse(CI=="(NA, NA)", NA, CI)) %>%
  dplyr::mutate(sign=ifelse(sign==1, "$\\star$", "")) %>%
  dplyr::relocate(CI, .before=sign)
## Make table
vect.align <- c(rep("l", 1), rep("r", 4), "c")
col.names <- c("Country -- study area", "fc2010", "Mean", "Sd", "CI 95%",
               "signif")
unit.vect <- c("", "(Kha)", "", "", "", "")
knitr::kable(df, caption="(ref:cap-road)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         digits=3,
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=TRUE, align=vect.align) %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=11, repeat_header_method="replace")
```

\newpage

<!--------------------------------->
<!-- Back-transformed parameters -->
<!--------------------------------->

## Tables S8--S9 -- Back-transformed parameters {-}

(ref:cap-bt-par) **Back-transformed parameters for each study area**. We back-transformed the parameters using the mean and standard-deviation of each continuous variable for each study area. Doing so, we can use Eq. \@ref(eq:icar) to compute the change in the probability of deforestation associated to a particular change in the explanatory variables, in their original units. To use this table of parameters, distances and elevation must be expressed in kilometers (Km), and slope must be expressed in hecto-degrees ($10^2$°). Note that the intercept is affected by the back-transformation but that the effect associated to protected areas ("pa") and the variance of the spatial random effects ("Vrho") are left unchanged.\vspace{0.5cm}

```{r bt-par, results="asis", echo=FALSE}
options(knitr.kable.NA="--")
## Import data
df <- read.table("tables/backtransformed_parameters.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::select(-area_cont, -area_ctry, -area_name)
## Make table
vect.align <- c(rep("l", 1), rep("r", 10))
col.names <- c("study area", "int", "pa", "elev", "slope",
               "ddefor", "dedge", "driver", "droad", "dtown", "Vrho")
unit.vect <- c("", "", "", "(Km)", "($10^2$°)", rep("(Km)", 5), "")
knitr::kable(df, caption="(ref:cap-bt-par)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         digits=3,
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=FALSE, align=vect.align) %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=10, repeat_header_method="replace")
```

(ref:cap-bt-par-reg) **Back-transformed parameters per region and continent**. We back-transformed the parameters using the mean and standard-deviation of each continuous variable for each study area. We then used the forest cover in 2010 to compute the back-transformed parameter estimate weighted mean per region. Doing so, we can use Eq. \@ref(eq:icar) to compute the change in the probability of deforestation associated with a particular change in the explanatory variables, in their original units. To use this table of parameters, distances and elevation must be expressed in kilometers (Km), and slope must be expressed in hecto-degrees ($10^2$°).\vspace{0.5cm}

```{r bt-par-region, results="asis", echo=FALSE}
options(knitr.kable.NA="--")
## Import data
df <- read.table("tables/backtransformed_weighted_param_region.csv", header=TRUE, sep=",")
## Make table
vect.align <- c(rep("l", 1), rep("r", 10))
col.names <- c("Region", "int", "pa", "elev", "slope",
               "ddefor", "dedge", "driver", "droad", "dtown", "Vrho")
unit.vect <- c("", "", "", "(Km)", "($10^2$°)", rep("(Km)", 5), "")
knitr::kable(df, caption="(ref:cap-bt-par-reg)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         digits=3,
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=FALSE, align=vect.align) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=10, repeat_header_method="replace")
```

\newpage

<!------------------------------------------------>
<!-- Mathematical formulas for accuracy indices -->
<!------------------------------------------------>

## Tables S10--S11 -- Mathematical formulas for accuracy indices {-}

(ref:cap-confusion-matrix) **Confusion matrix used to compute accuracy indices**. A confusion matrix can be computed to compare model predictions with observations.\vspace{0.5cm}

```{r confusion-matrix, results="asis", echo=FALSE}
df <- read.table("tables/confusion-matrix.txt", header=TRUE, sep=",")
vect.align <- rep("l",2)
col.names <- c("","","Observations","","Total")
options(knitr.kable.NA="")
knitr::kable(df, caption="(ref:cap-confusion-matrix)", booktabs=TRUE,
                         escape=FALSE,
                         col.names=col.names,
                         align=vect.align, linesep="") %>%
    kable_styling(latex_options=c("HOLD_position","striped"),
                  full_width=FALSE, position="center")
```


(ref:cap-accuracy-indices) **Formulas used to compute accuracy indices**. Several accuracy indices can be computed from the confusion matrix to estimate and compare models' predictive performance. We followed the definitions of @Pontius2008 for the FOM and @Liu2011 for the other indices. Note that the AUC relies on the predicted probabilities for observations 0 (non-deforested) and 1 (deforested), not on the confusion matrix.\vspace{0.5cm}

```{r accuracy-indices, results="asis", echo=FALSE}
df <- read.table("tables/accuracy-indices.txt", header=TRUE, sep=",")
# Remove some indices
df <- df %>%
  dplyr::filter(!(Index %in% c("Expected Accuracy", "Cohen's Kappa")))
vect.align <- rep("l",2)
col.names <- c("Index","Formula")
knitr::kable(df, caption="(ref:cap-accuracy-indices)", booktabs=TRUE,
                         escape=FALSE,
                         col.names=col.names,
                         align=vect.align, linesep="") %>%
    kable_styling(latex_options=c("HOLD_position","striped"), full_width=FALSE, position="center")
```

<!---------------------->
<!-- Accuracy indices -->
<!---------------------->

## Tables S12--S13 -- Accuracy indices {-}

(ref:cap-accuracy) **Accuracy indices' weighted means for the three statistical models**. Accuracy indices were averaged across study areas using forest cover areas in 2010 as weights. D: percentage of deviance explained, AUC: Area Under ROC Curve, OA: overall accuracy, FOM: Figure Of Merit, TSS: True Skill Statistics. Averaged accuracy indices were computed for the three statistical models: "glm", "icar", and "rf" model. While the "rf" model has a higher percentage of deviance explained in average (higher goodness-of-fit), the "icar" model has higher values of accuracy indices from the cross-validation (higher predictive power).\vspace{0.5cm}

```{r accuracy, results="asis", echo=FALSE}
df <- read.table("tables/perf_mod.csv", header=TRUE, sep=",")
vect.align <- c(rep("l", 1), rep("r", 5))
col.names <- c("Model","D","AUC","OA","FOM","TSS")
knitr::kable(df, caption="(ref:cap-accuracy)", booktabs=TRUE,
                         escape=FALSE,
                         digits=1,
                         col.names=col.names,
                         align=vect.align, linesep="") %>%
    kable_styling(latex_options=c("HOLD_position","striped"), stripe_index=c(2), full_width=FALSE, position="center")
```


(ref:cap-accuracy-cont) **Accuracy indices' weighted means for the three statistical models by continent**. Accuracy indices were averaged across study areas for each continent using forest cover areas in 2010 as weights. D: percentage of deviance explained, AUC: Area Under ROC Curve, OA: overall accuracy, FOM: Figure Of Merit, TSS: True Skill Statistics. Averaged accuracy indices were computed for the three statistical models: "glm", "icar", and "rf" model.\vspace{0.5cm}

```{r accuracy-cont, results="asis", echo=FALSE}
df <- read.table("tables/perf_cont_mod.csv", header=TRUE, sep=",")
vect.align <- c(rep("l", 1), rep("r", 5))
col.names <- c("Continent","Model","D","AUC","OA","FOM","TSS")
knitr::kable(df, caption="(ref:cap-accuracy-cont)", booktabs=TRUE,
                         escape=FALSE,
                         digits=1,
                         col.names=col.names,
                         align=vect.align, linesep="") %>%
  column_spec(1, bold=TRUE) %>%
  collapse_rows(columns=1, latex_hline = "major", valign = "middle") %>%
  kable_styling(latex_options=c("HOLD_position","striped"), stripe_index=c(2,5,8), full_width=FALSE, position="center")
```

\newpage

<!------------------------------------>
<!-- Past forest cover change -->
<!------------------------------------>

## Tables S14--S15 -- Past forest cover change {-}

(ref:cap-fcc-hist) **Past forest cover change for each study area**. Forest cover areas are given in thousand hectares (Kha) for the years 2000, 2010 and 2020 ("fc2000", "fc2010", and "fc2020", respectively). The mean annual deforested area $d$ for the ten-year period 2010--2020 is given in hectare per year (ha/yr). The corresponding mean annual deforestation rate $p$ is also provided in percent per year (%/yr), with one decimal precision, to be able to compare the intensity of deforestation between study areas.\vspace{0.5cm}

```{r fcc-hist, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/forest_cover_change_mean.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::select(area_name, for2000, for2010, for2020, andef, pdef) %>%
  mutate_at(.var=vars(starts_with("for")),.fun=function(x){round(x/1000)})
## Make table
vect.align <- c(rep("l", 1), rep("r", 5))
col.names <- c("Country -- study area", "fc2000", "fc2010", "fc2020", "$d$", "$p$")
unit.vect <- c("", "(Kha)", "(Kha)", "(Kha)", "(ha/yr)", "(\\%/yr)")
knitr::kable(df, caption="(ref:cap-fcc-hist)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=FALSE, align=vect.align) %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=11, repeat_header_method="replace")
```

(ref:cap-fcc-hist-reg) **Past forest cover change per region and continent**. Areas of forest cover are given in thousand hectares (Kha) for the years 2000, 2010 and 2020 ("fc2000", "fc2010", and "fc2020", respectively). The mean annual deforested area $d$ for the ten-year period 2010--2020 is given in hectare per year (ha/yr). The corresponding mean annual deforestation rate $p$ is also provided in percent per year (%/yr), with one decimal precision, to be able to compare the intensity of deforestation between study areas. Estimates for America include Brazil, and estimates for Asia include India. Around 6.4 Mha (64,000 km$^2$, about half the size of Greece or the size of West Virginia) of natural old-growth moist tropical forest have been disappearing each year in the period 2010--2020.\vspace{0.5cm}

```{r fcc-hist-reg, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/fcc_hist_region_mean.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::filter(!(area_cont %in% c("DRC", "Indonesia"))) %>%
  dplyr::select(area_cont, for2000, for2010, for2020, andef, pdef) %>%
  mutate_at(.var=vars(starts_with("for")),.fun=function(x){round(x/1000)})
## Make table
vect.align <- c(rep("l", 1), rep("r", 5))
col.names <- c("Region", "fc2000", "fc2010", "fc2020", "$d$", "$p$")
unit.vect <- c("", "(Kha)", "(Kha)", "(Kha)", "(ha/yr)", "(\\%/yr)")
knitr::kable(df, caption="(ref:cap-fcc-hist-reg)", booktabs=TRUE, longtable=FALSE,
                         format.args=list(big.mark=","),
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=FALSE, align=vect.align) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=11, repeat_header_method="replace")
```

\newpage

<!------------------------------>
<!-- Forest cover projections -->
<!------------------------------>

## Tables S16--17 -- Forest cover projections {-}

(ref:cap-fcc-proj) **Forest cover projections for each study area**. Projected areas of forest cover are given in thousand hectares (Kha) for four years in the future (2040, 2060, 2080, and 2100). Projections were made using the forest cover in 2020 and the mean annual deforested area in the ten-year period 2010--2020 ("fc2000" and $d$ respectively in Table \@ref(tab:fcc-hist)), assuming a "business-as-usual" scenario of deforestation. Column "loss21" indicates the projected percentage of forest cover loss during the 21$^\text{st}$ century (2100 vs. 2000). Column "yrdis" indicates the estimated year at which all the forest of the study area will have disappeared.\vspace{0.5cm}

```{r fcc-proj, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/forest_cover_change_mean.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::mutate(yrdis=as.character(yrdis)) %>%
  dplyr::select(area_name, for2040, for2060, for2080, for2100, loss21, yrdis) %>%
  mutate_at(.var=vars(starts_with("for")),.fun=function(x){round(x/1000)})
## Make table
vect.align <- c(rep("l", 1), rep("r", 6))
col.names <- c("Country -- study area", "fc2040", "fc2060", "fc2080", "fc2100", "loss21", "yrdis")
unit.vect <- c("", "(Kha)", "(Kha)", "(Kha)", "(Kha)", "(\\%)", "")
knitr::kable(df, caption="(ref:cap-fcc-proj)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=FALSE, align=vect.align) %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=11, repeat_header_method="replace")
```

(ref:cap-fcc-proj-reg) **Forest cover projections per region and continent**. Projected areas of forest cover are given in thousand hectares (Kha) for four dates in the future (2040, 2060, 2080, and 2100). Projections were made using the forest cover in 2020 and the mean annual deforested area in the ten-year period 2010--2020 ("fc2000" and $d$ respectively in Table \@ref(tab:fcc-hist-reg)), assuming a "business-as-usual" scenario of deforestation. Column "loss21" indicates the projected percentage of forest cover loss during the 21$^\text{st}$ century (2100 vs. 2000). At the continental level, it makes less sense to compute the year at which all the forest will have disappeared, as some countries might conserve forest for a very long time, even though they account for a very small proportion of the total forest area at the continental scale. Instead, we computed the estimated year at which 75% of the forest cover in 2000 will have disappeared ("yr75dis").\vspace{0.5cm}

```{r fcc-proj-reg, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/fcc_proj_region_mean.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::filter(!(cont %in% c("DRC", "Indonesia"))) %>%
  dplyr::mutate(yr75dis=as.character(yr75dis)) %>%
  dplyr::select(cont, for2040, for2060, for2080, for2100, loss21, yr75dis) %>%
  mutate_at(.var=vars(starts_with("for")),.fun=function(x){round(x/1000)})
## Make table
vect.align <- c(rep("l", 1), rep("r", 6))
col.names <- c("Region", "fc2040", "fc2060", "fc2080", "fc2100", "loss21", "yr75dis")
unit.vect <- c("", "(Kha)", "(Kha)", "(Kha)", "(Kha)", "(\\%)", "")
knitr::kable(df, caption="(ref:cap-fcc-proj-reg)", booktabs=TRUE, longtable=FALSE,
                         format.args=list(big.mark=","),
                         digits=0,
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=FALSE, align=vect.align) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=11, repeat_header_method="replace")
```

<!------------------------------------------------->
<!-- Endemic species in biodiversity hotspots    -->
<!------------------------------------------------->

## Table S18 -- Endemic species in the biodiversity hotspots {-}

(ref:cap-hotspots) **Plant and vertebrate species endemic to each of the biodiversity hotspots**. Estimates come from @Mittermeier2011. P: Plants, B: Birds, R: Reptiles, F: Freshwater fishes, A: Amphibians, M: Mammals.\vspace{0.5cm}

```{r hotspots, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/species_biodiversity_hotspots.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::select(Hotspot, ends_with("_E"))
## Make table
vect.align <- c(rep("l", 1), rep("r", 6))
col.names <- c("Hotspot", "P", "B", "R", "F", "A", "M")
knitr::kable(df, caption="(ref:cap-hotspots)", booktabs=TRUE, longtable=FALSE,
                         format.args=list(big.mark=","),
                         digits=0,
                         escape=FALSE,
                         col.names=col.names,
                         align=vect.align, linesep="") %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=11, repeat_header_method="replace")
```

\newpage

<!------------------------------------------------->
<!-- Carbon emissions associated with deforestation -->
<!------------------------------------------------->

## Table S19 -- Cumulative carbon emissions associated with deforestation {-}

(ref:cap-c-em-sm) **Cumulative carbon emissions associated with future deforestation**. We computed the cumulative carbon emissions associated with future deforestation from 2020 for each study area (C in Gg, $10^9$ g). To do so, we used our maps of projected forest cover change together with available global or pantropical maps of aboveground biomass (either the WUR, WHRC or CCI map). We present here the results obtained with the ESA CCI aboveground biomass map by @Santoro2021.\vspace{0.5cm}

```{r c-em-sm, results="asis", echo=FALSE}
options(knitr.kable.NA="--")
## Import data
df <- read.table("tables/C_emissions_mean_cci.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::select(area_name, C2040, C2060, C2080, C2100) %>%
  mutate_at(.var=vars(starts_with("C")), ~na_if(., 0)) %>%
  mutate_at(.var=vars(starts_with("C")),.fun=function(x){round(x/1000)})
 
## Make table
vect.align <- c(rep("l", 1), rep("r", 4))
col.names <- c("Country -- study area", "C2040", "C2060", "C2080", "C2100")
unit.vect <- c("", "(Gg)", "(Gg)", "(Gg)", "(Gg)")
knitr::kable(df, caption="(ref:cap-c-em-sm)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=FALSE, align=vect.align) %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=11, repeat_header_method="replace")
```

\newpage

<!--------------------------------------------->
<!-- Uncertainty -->
<!--------------------------------------------->

## Table S20 -- Confidence interval of the annual deforested area {-}

(ref:cap-ci-d) **Confidence interval of the annual deforested area**. We computed the 95% confidence interval of the annual deforested area $d$ (in ha/yr) for each study area. We used the deforestation observations $d_t$ for the ten years $t$ from 2010 to 2020. The lower and upper bounds of the confidence interval were denoted $d'$ and $d''$, respectively. Forest cover areas (in Kha) for the years 2010 and 2020 ("fc2010" and "fc2020") are also provided for comparison.\vspace{0.5cm}

```{r ci-d, results="asis", echo=FALSE}
## Import data
df <- read.table("tables/forest_cover_change_mean.csv", header=TRUE, sep=",")
## Arrange data
df <- df %>%
  dplyr::select(area_name, for2010, for2020, d_mean, d_min, d_max) %>%
  mutate_at(.var=vars(starts_with("for")),.fun=function(x){round(x/1000)})
## Make table
vect.align <- c(rep("l", 1), rep("r", 5))
col.names <- c("Country -- study area", "fc2010", "fc2020", "$d$", "$d'$", "$d''$")
unit.vect <- c("", "(Kha)", "(Kha)", "(ha/yr)", "(ha/yr)", "(ha/yr)")
knitr::kable(df, caption="(ref:cap-ci-d)", booktabs=TRUE, longtable=TRUE,
                         format.args=list(big.mark=","),
                         escape=FALSE,
                         col.names=unit.vect,
                         align=vect.align, linesep="") %>%
  add_header_above(header=col.names, line=FALSE, escape=FALSE, align=vect.align) %>%
  pack_rows("America", 1, 64) %>%
  pack_rows("Africa", 65, 96) %>%
  pack_rows("Asia", 97, 119) %>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),
                full_width=FALSE,
                position="center", font_size=11, repeat_header_method="replace")
```

\newpage

# Legends for Supplementary Data

Supplementary Data are available at <https://forestatrisk.cirad.fr/data-s.html>.

## Data S1 -- Uncertainty around projected forest cover {-}

Data S1: **Uncertainty around projected forest cover**. Past and projected forest cover change by study area are given in thousand hectares (Kha). The mean annual deforested area $d$ for the ten-year period 2010--2020 is given in hectare per year (ha/yr). The corresponding mean annual deforestation rate $p$ is also provided in percent per year (%/yr), with one decimal precision, to be able to compare the intensity of deforestation between study areas. Projections were made using the forest cover in 2020 ("fc2000") and either (i) the mean annual deforested area $d$ for the ten-year period 2010--2020, (ii) the lower bound $d'$ of the confidence interval for the annual deforested area, or (iii) the upper bound $d''$ of the confidence interval for the annual deforested area. We considered a business-as-usual scenario of deforestation (deforestation constant through time) for the projections. Column "loss21" indicates the projected percentage of forest cover loss during the 21^st^ century (2100 vs. 2000). Column "yrdis" indicates the estimated year at which all the forest of the study area will have disappeared.

## Data S2 -- Uncertainty around projected carbon emissions {-}

Data S2: **Uncertainty around projected carbon emissions**. We combined our maps of projected forest cover change together with the aboveground biomass map by @Avitabile2016 to compute the cumulative carbon emissions associated with future deforestation from 2020 for each study area (C in Gg=10^9^ g). Maps of projected forest cover change were derived using either (i) the mean annual deforested area $d$ for the ten-year period 2010--2020, (ii) the lower bound $d'$ of the confidence interval for the annual deforested area, or (iii) the upper bound $d''$ of the confidence interval for the annual deforested area. Column "C2020" indicates the carbon emissions associated with past deforestation during the period 2010--2020 for comparison.

\newpage
