---
title: "Rasters"
---

Raster maps from this study are available as Cloud Optimized GeoTIFF ([COG](https://www.cogeo.org/)) served on the ForestAtRisk website or on Google Earth Engine ([GEE](https://earthengine.google.com/)). 

### Cloud Optimized GeoTIFF

- [List of all rasters](https://forestatrisk.cirad.fr/tropics/tif/)
- [COG tutorial](https://forestatrisk.cirad.fr/notebooks/cog.html)

#### _Table of rasters (some)_

```{r cogs, echo=FALSE, results='asis'}
# Create data frame
raster_type <- c("Past deforestation 2000-2010-2020",
				 "Deforestation probability in 2020",
				 "Projected deforestation 2020-2050",
				 "Projected deforestation 2020-2100")
df <- data.frame(rast=raster_type, America=NA, Africa=NA, Asia=NA)
cont <- c("America", "Africa", "Asia")
cont_ab <- c("AME", "AFR", "ASI") # Abbreviation
url_base <- "https://forestatrisk.cirad.fr/tropics/tif/"

# Loop on continents
for (i in 1:length(cont)) {
	# Continent abbreviation
	cab <- cont_ab[i]
	# fcc123
	text_fcc <- glue("fcc123_{cab}.tif")
	url_fcc <- glue("{url_base}{text_fcc}")	
	# Prob
	text_prob <- glue("prob_{cab}.tif")
	url_prob <- glue("{url_base}{text_prob}")
	# Projected 2050
	text_fcc_2050 <- glue("fcc_2050_{cab}.tif")
	url_fcc_2050 <- glue("{url_base}{text_fcc_2050}")
	# Projected 2100
	text_fcc_2100 <- glue("fcc_2100_{cab}.tif")
	url_fcc_2100 <- glue("{url_base}{text_fcc_2100}")
	# 
	df[cont[i]] <- c(glue("<a href='{url_fcc}'>{text_fcc}</a>"),
					 glue("<a href='{url_prob}'>{text_prob}</a>"),
					 glue("<a href='{url_fcc_2050}'>{text_fcc_2050}</a>"),
					 glue("<a href='{url_fcc_2100}'>{text_fcc_2100}</a>"))
}

# Table
col_names <- c("", "America", "Africa", "Asia")
vect_align <- c("l", rep("c", 3))
knitr::kable(df, booktabs=TRUE, longtable=FALSE,
			 format.args=list(big.mark=","),
			 escape=FALSE,
			 col.names=col_names,
			 align=vect_align, linesep="") %>%
	kable_styling(full_width=FALSE,
	              bootstrap_options=c("striped", "hover"),
	              position="center")
```

#### _Rasters resolution and projection_

Rasters have a resolution of 30 m and are available in Albers equal area conic projections (one different projection for each continent):

- America: [South America Albers Equal Area Conic](https://epsg.io/102033)
- Africa: [Africa Albers Equal Area Conic](https://epsg.io/102022)
- Asia: [Asia South Albers Equal Area Conic](https://epsg.io/102028)

### Google Earth Engine

We provide a [JavaScript code](https://code.earthengine.google.com/d10bacab3bf0994097c37678e8098e7d) for importing and using the ForestAtRisk assets in Google Earth Engine.

```{r assets, echo=FALSE, results='asis'}
# Create data frame
asset <- c("Past deforestation 2000-2010-2020",
           "Deforestation probability in 2020",
           "Projected deforestation 2020-2050",
           "Projected deforestation 2020-2100")
asset_id <- c("projects/forestatrisk/assets/v1_2020/fcc_123",
              "projects/forestatrisk/assets/v1_2020/prob_2020",
              "projects/forestatrisk/assets/v1_2020/fcc_2050",
              "projects/forestatrisk/assets/v1_2020/fcc_2100")
df <- data.frame(asset=asset, asset_id=asset_id)

# Table
col_names <- c("Asset", "Asset ID")
vect_align <- c("ll")
knitr::kable(df, booktabs=TRUE, longtable=FALSE,
			 format.args=list(big.mark=","),
			 escape=FALSE,
			 col.names=col_names,
			 align=vect_align, linesep="") %>%
	kable_styling(full_width=FALSE,
	              bootstrap_options=c("striped", "hover"),
	              position="center")
```
