
Tropical forests are at the heart of concerns when it comes to climate change and biodiversity loss, which both pose unprecedented threats to human civilization [@IPCC2014; @Cardinale2012]. Commonly called the "Jewels of the Earth", tropical forests shelter 30 million species of plants and animals, representing half of the Earth's wildlife and at least two-thirds of its plant species [@Gibson2011; @Wilson2012]. Through carbon sequestration, they play an important role in the global carbon cycle and regulate the global climate [@Baccini2017]. Intact tropical forest ecosystems also prevent outbreaks of zoonoses and reduce the risk of pandemics [@Tollefson2020]. At local scale, tropical forests regulate the regional climate [@Dickinson1992], cooling the atmosphere [@Ellison2017], and facilitate access to water [@Ellison2017; @Zhang2021]. They also protect against erosion and flooding [@Bradshaw2007]. Close to 1.6 billion people (a quarter of the world's population) rely on forest resources for their livelihoods [@FAO2015]. Despite the many ecosystem services they provide, tropical forests are disappearing at an alarming rate [@FAO2015; @Hansen2013; @Vancutsem2021; @Achard2014], mostly because of human activities [@Curtis2018; @Geist2002]. Past studies have estimated that about 8 Mha of tropical forest (including moist and dry forests) are currently disappearing each year [@FAO2015]. At this rate, will there still be any tropical forests left at the end of the 21^st^ century, if so, where will they be concentrated, and what will be the consequences of tropical deforestation on climate and biodiversity in the future?

Forecasting forest cover change is paramount as it allows us to foresee the consequences of deforestation (in terms of carbon emissions, biodiversity loss, or water supply) under various technological, political, and socio-economic scenarios, and to inform decision makers accordingly [@Clark2001]. The "business-as-usual" scenario (i.e., in the absence of any change in deforestation in the future) is of particular interest as it makes it possible to predict the likely future in the absence of change, and if necessary, to alert decision-makers to an essential change of course to avoid any potential environmental disaster. Although carbon dioxide emission and climate change scenarios have been developed for several years by the Intergovernmental Panel on Climate Change [@IPCC2014] and are now widely used by the scientific community and known to the general public, no equivalent scenarios for land-use change and biodiversity at the global scale are currently available. Spatialized forest cover change scenarios are crucial because both forest carbon stocks [@Baccini2017; @Avitabile2016] and biodiversity [@Kremen2008; @Mittermeier2011] vary considerably in space at fine scale. Non-spatial scenarios of forest cover change [@FAO2015] cannot be used to forecast associated carbon emissions and change in biodiversity accurately, or for systematic conservation planning at the local scale. Spatial forecasts of forest cover change are based on spatial statistical models, which enable estimation of a probability of change in space as a function of a set of spatial predictors. In addition to forecasts, statistical models can be used to identify the main drivers of deforestation and to quantify their relative effects. For example, models can be used to assess the impact of roads on the risk of deforestation [@Laurance2014] and the effectiveness of protected areas at reducing deforestation [@Andam2008; @Wolf2021].

\newpage

Few authors have attempted to provide spatialized forest cover change scenarios in the tropics at large spatial scales. The largest studies to date have focused on modelling and forecasting forest cover change at the scale of the Amazonian basin [@Soares-Filho2006]. In this paper, we present the first attempt to model and forecast tropical deforestation at the pantropical scale using high-resolution spatial data. This was made possible by the recent availability of pantropical spatial datasets of forest cover change [@Vancutsem2021] and of global spatial datasets of explanatory factors related to deforestation at the required resolution (World Database on Protected Areas, SRTM Digital Elevation Database, and OpenStreetMap). We combine these extensive datasets in a spatial statistical model to test the effectiveness of protected areas at reducing deforestation and to assess the impact of roads on the risk of deforestation. Assuming a business-as-usual scenario, we derive high-resolution maps of deforestation risk and future forest cover over the 21^st^ century in the humid tropics. We also estimate the carbon emissions associated with projected deforestation and conduct an uncertainty analysis.

# Remaining tropical moist forests in 2100

```{r fcc-proj, results="asis", echo=FALSE}
require(dplyr)
require(here)
## Import data
df <- read.table(here::here("Manuscript/Supplementary_Materials/tables/forest_cover_change_mean.csv"), header=TRUE, sep=",")
## All study areas
nctry_loss <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::group_by(area_cont) %>%
  dplyr::summarize(nctry=n(),
                   nctry_loss21_100=sum(loss21==100),
                   nctry_loss21_75=sum(loss21>=75))
## Study-areas with more than 1 Mha of forest in 2020
nctry_loss_large <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::filter(for2020>=1e6) %>%
  dplyr::group_by(area_cont) %>%
  dplyr::summarize(nctry=n(),
                   nctry_loss21_100=sum(loss21==100),
                   nctry_loss21_75=sum(loss21>=75))
## List countries
## Study-areas with more than 1 Mha of forest in 2020
ctry <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::filter(for2020>=1e6 & loss21==100) %>%
  dplyr::arrange(area_cont, -for2020)
```

Using Vancutsem _et al._'s study [@Vancutsem2021] as a reference, we estimate that around 7.6 Mha (76,000 km^2^, about the size of Scotland or South Carolina) of tropical moist forest have been disappearing each year over the last decade (2010--2020). We show here that under a business-as-usual scenario of deforestation, 61% of the world's tropical moist forest will disappear over the course of the 21^st^ century (Table \@ref(tab:fcc)) but with marked differences in the percentage of forest cover loss at the continental (Table \@ref(tab:fcc) and Fig. S12) and country (Tables S14 and S16) scales. In Southeast Asia, where the forest area remaining in 2020 was relatively low (247 Mha) and the area deforested each year is relatively large (2.5 Mha/yr) compared to the two other continents, the percentage of forest cover loss over the 21^st^ century will reach 78%. In Africa, where the annual deforested area is lower (2.2 Mha/yr), this percentage will be 58%. In Latin America, where the annual deforested area is higher (2.9 Mha/yr), but where the remaining tropical moist forest in 2020 is also much larger than in Southeast Asia and Africa (621 Mha), this percentage will be 41%. Under a business-as-usual scenario of deforestation, three quarters of the tropical moist forests that existed in 2000 will have disappeared around years 2100, 2150, and 2200 in Southeast Asia, Africa, and Latin America, respectively (Table \@ref(tab:fcc) and Fig. S12).

At the country scale, we predict that 47 countries (18 in Latin America, 21 in Africa, and 8 in Southeast Asia) out of the 92 we studied, plus 18 states in Brazil and one region in India, will lose all their tropical forests by 2100 (Fig. \@ref(fig:fcc2100) and Table S16). Among these countries or regions, 20 countries (6 in America, 8 in Africa, and 6 in Asia), five states in Brazil, and one region in India had more than one million hectares of forest in 2020. These 20 countries include most of the countries of Central America (Mexico, Guatemala, Honduras, Nicaragua), Western (Sierra Leone, Ivory Coast, Ghana, Nigeria) and Eastern (Ethiopia, Madagascar) Africa, and continental Southeast Asia (Myanmar, Laos, Vietnam, Cambodia, Malaysia, Indian Western Ghats). These regions include almost all tropical forests located within six biodiversity hotspots identified by Mittermeier _et al._ [@Mittermeier2011]: Mesoamerica, Guinean Forests of West Africa, the Horn of Africa, Madagascar and Indian Ocean Islands, Indo-Burma, Western Ghats and Sri Lanka (Fig. \@ref(fig:fcc2100)).

Using an advanced spatial statistical modelling approach (see Methods), we obtain high resolution (30 m) pantropical maps of the deforestation risk (Fig. \@ref(fig:prob)) and project forest cover for the 21^st^ century in the humid tropics (Fig. \@ref(fig:fcc2100)) under a business-as-usual scenario. Three large "blocks" of relatively intact tropical moist forest will remain in 2100 (see both Fig. \@ref(fig:fcc2100) and the interactive map of projected deforestation for the period 2020--2100 available at <https://forestatrisk.cirad.fr/maps.html>). One forest block will be located in Latin America and will include the upper part of the Amazonian basin (including the Peruvian, Ecuadorian, Colombian and Venezuelan Amazonia) and the Guiana Shield (Guyana, Suriname, and French Guiana). The second block will be located in the western part of the Congo basin and will include forests in Gabon, Equatorial Guinea, Cameroon, the Central African Republic, and the Republic of Congo. The third block will be located in Melanesia and will include forests in Papua New Guinea, Solomon Islands, and Vanuatu.

Apart from these three large and relatively intact forest blocks, the tropical moist forest remaining in 2100 will be highly fragmented (Fig. \@ref(fig:fcc2100)). In Latin America, highly fragmented forests will be found in the Brazilian states of the Amazonian deforestation arc (Acre, Rondonia, Mato Grosso, Para, Amapa) and in the Roraima state in the northern Amazonia. In Africa, forests in the Democratic Republic of the Congo (DRC) will also be highly fragmented and will be completely separated from the large forest block located in the western part of the Congo basin. In Southeast Asia, small patches of heavily fragmented forests will remain in Thailand, Indonesia, and the Philippines (Fig. \@ref(fig:fcc2100)). The remaining forests will be concentrated in remote areas (far from roads and towns), preferentially in protected areas, and at high elevations. For example, the remaining forests of Borneo will be concentrated in the Betung Kerihun and Kayan Mentarang National Parks.

As tropical forests shelter a large proportion of terrestrial biodiversity and carbon stocks on land, future tropical deforestation is expected to have strong negative impacts on both biodiversity and climate. The impact of projected deforestation on carbon emissions is discussed below, but rigorous assessment of the impact of projected deforestation on biodiversity is beyond the scope of this study. Such an impact analysis would require accurate species distribution and biodiversity maps including a large number of species representative of the biodiversity in the tropics. Such maps are not available to date [@Pimm2014]. Nonetheless, as a rough estimate, if we consider only endemic species [@Mittermeier2011] in the six biodiversity hotspots where almost all the tropical forest is predicted to disappear by 2100, and assume that most of these species depend on tropical moist forests, deforestation would lead to the extinction of 29,140 species of plants and 4,576 species of vertebrates (including birds, reptiles, amphibians, freshwater fishes, and mammals) which cannot be found anywhere else on Earth (Table S18).

\newpage

# Carbon emissions under a business-as-usual scenario of deforestation

Here we estimate the aboveground carbon emissions associated with deforestation projected for the period 2020--2100 under a business-as-usual scenario of deforestation. When computing carbon emissions associated with projected deforestation, we assume that the carbon stocks of existing forests will remain stable in the future. Under a business-as-usual scenario of deforestation (i.e., constant annual deforested area), the change in predicted annual carbon emissions is only attributable to the spatial distribution of forest carbon stocks and to the location of the future deforestation. We find that annual carbon emissions associated with deforestation of tropical moist forests will increase from 0.525 Pg/yr in 2010--2020 to 0.746 Pg/yr in 2080--2090, which corresponds to a 42% increase (Fig. \@ref(fig:c-em)). This increase in annual carbon emissions is predicted for all three continents (Fig. \@ref(fig:c-em)). A decrease in carbon emissions is then predicted starting from the period 2070--2080 for Southeast Asia and the period 2080--2090 at pantropical scale (Fig. \@ref(fig:c-em)).

The predicted increase in annual carbon emissions is explained by the fact that the forests which will be deforested in the future have higher carbon stocks. Several studies have shown that elevation is an important variable in determining forest carbon stocks [@Vieilledent2016; @Saatchi2011]. Forest carbon stocks are expected to be optimal at mid-elevation [@Vieilledent2016] due to higher orographic precipitation at this elevation and because the climatic stress associated with strong winds and low temperature is lower at mid-elevation than at high elevation. Here, we show that low-elevation areas are more deforested than high-elevation areas (Tables S4 and S5). This is explained by the fact that low-elevation areas are more accessible to human populations and by the fact that arable lands are concentrated at low elevation, where the terrain slope is usually lower and the soil is more productive [@Geist2002]. Consequently, the predicted increase in carbon emissions can be explained by the fact that deforestation will move towards higher elevation areas where forest carbon stocks are higher. Moreover, remote forest areas that have been less disturbed by human activities in the past have accumulated large quantities of carbon [@Dargie2017; @Brinck2017]. The progressive deforestation of more intact forests also explains the predicted increase in carbon emissions.

The decrease in carbon emissions predicted from the period 2070--2080 for Southeast Asia, and from the period 2080--2090 at pantropical scale, can be associated with a decrease in carbon stocks of deforested areas (in association with environmental factors, such as lower carbon stocks at very high elevation) or a decrease in the total deforested area at the continental and global scale, as countries progressively lose all their forest. In Southeast Asia, six countries will lose all their forest between 2070 and 2100 (Table S16). These countries (which include Laos, Malaysia, Myanmar, and Vietnam) account for a significant proportion (32%) of the annual deforested area in Southeast Asia (808,375 ha/yr out of 2,532,927 ha/yr, see Tables S14 and S15). This largely explains the predicted decrease in carbon emissions in Southeast Asia from 2070 on.

Our estimates of 0.525 Pg/yr (0.208, 0.199, and 0.118 Pg/yr for Latin America, Southeast Asia, and Africa, respectively) of aboveground carbon emissions due to tropical deforestation for the period 2010--2020 are consistent with those of previous studies [@Baccini2017; @Achard2014; @Harris2012]. For the period 2000--2014, previous studies estimated 0.81--0.88 Pg/yr of carbon emissions associated with deforestation considering both moist and dry tropical forests, while our study only focuses on tropical _moist_ forests. Because of the projected 42% increase in carbon emissions associated with tropical deforestation under a business-as-usual scenario, tropical forests will act as an increasing net carbon source [@Baccini2017], thus reinforcing climate change in the future.

These results demonstrate the importance of spatial predictions of deforestation to forecast carbon emissions associated with future deforestation. Using mean annual forest cover change estimates per country and forest type (such as those provided by the Forest Resource Assessment report [@FAO2015]) and mean forest carbon stocks per continent and forest type (such as the emission factors provided by IPCC [@IPCC2019]), it is not possible to predict future trends in carbon emissions associated with deforestation. Under a business-as-usual scenario of deforestation, projected carbon emissions will result from the combination of the spatial variation in forest carbon stocks and the location of the future deforestation.

# Effectiveness of protected areas at displacing deforestation

```{r pa-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="PA"]
nsignif <- df$nctry_sign[df$var=="PA"]
perc <- df$perc[df$var=="PA"]
perc_w <- df$perc_w[df$var=="PA"]
```

```{r pa-decrease, results="asis", echo=FALSE}
## Inv_logit function
inv_logit <- function (x, min=0, max=1) {
    p <- exp(x)/(1 + exp(x))
    p <- ifelse(is.na(p) & !is.na(x), 1, p)
    p * (max - min) + min
}
## Effect of protected areas at decreasing deforestation probability
alpha_normalized <- -2.194
coef_pa_km <- -0.483 # Back-transformed parameter to have slope in km^-1
theta_mean <- inv_logit(alpha_normalized) # Mean deforestation probability
theta_pa <- inv_logit(alpha_normalized + coef_pa_km)
d_pa <- 100*round(1-(theta_pa/theta_mean), 2)
```

Forested protected areas are usually located in areas of high or unique biodiversity, in a non-random way. They are also usually found in remote places with less human disturbances and reduced accessibility, i.e. far from roads or cities (Fig. S6), and usually at higher elevation, because low-lying arable lands have already been preempted for agriculture [@Geist2002]. As a consequence, it is usually difficult to unravel the effect of protected areas from other correlated variables, for example distance to the nearest road, city, or elevation [@Andam2008]. The multivariate logistic regression model we use makes it possible to disentangle the effect of each explanatory variable in the spatial deforestation process. Moreover, the spatial random effects included in our model correct the potential bias in the protected area effect that could be associated with other unmeasured confounding variables, such as population density [@Andam2008].

Here we show that protected areas significantly reduce the risk of deforestation in `r nsignif` study areas out of `r nsa` (`r perc`% of the study areas). These `r nsignif` study areas accounted for `r perc_w`% of the tropical moist forest in 2010 (Table S6). On average, protected areas reduce the risk of deforestation by `r d_pa`% (Fig. \@ref(fig:proba-var) and Tables S5 and S9). This result clearly demonstrates the efficiency of protected areas at reducing the spatial risk of deforestation in the tropics. In a recent global study, Wolf _et al._ [@Wolf2021] found that protected areas reduced deforestation rates by 41%, close to the `r d_pa`% we find here by focusing on tropical moist forests. While past studies have assessed the effect of protected areas at reducing deforestation in particular countries or regions [@Andam2008; @Bruner2001] or at efficiently protecting a particular group of species [@Cazalis2020], our study is one of the first (see also [@Wolf2021; @Yang2021]) to assess the effectiveness of protected areas at reducing deforestation at the pantropical scale. Previous global studies [@Wolf2021; @Yang2021] were at 1 km resolution and used spatial matching methods and tree cover loss data [@Hansen2013]. Our approach is based on more accurate forest cover change maps in the humid tropics, in particular in Africa [@Vancutsem2021], and accounts for fine scale deforestation factors acting at a much smaller distance than the distance imposed by a 1 km resolution (see the effect of the distance to forest edge discussed below). Moreover, contrary to most spatial matching methods [@Andam2008; @Schleicher2019], the statistical model we use allows us to account for any potential confounding variables which might skew the estimated effect of protected areas.

Like other studies reporting the effect of protected areas on deforestation, our study demonstrates that protected areas are effective at _displacing_ deforestation outside protected areas in tropical countries, but not necessarily that protected areas play a role in _reducing_ the deforestation intensity per se. Indeed, the factors that drive the intensity of deforestation at the country scale are more socio-economic or political, such as the level of economic development, which determines people's livelihood and the link between people and deforestation [@Geist2002], the size of the population [@Barnes1990], or the environmental policy [@Soares-Filho2014]. In tropical countries with weak governance (where environmental law enforcement is low) and with a low level of development (where the pressure on forest is high), it is very unlikely that protected areas will remain forested. Under a business-as-usual scenario of deforestation, we assume that the deforestation intensity will remain constant over time. When all the forest outside the protected areas is deforested, deforestation is expected to occur inside protected areas (Fig. \@ref(fig:fcc2100)). In this scenario, protected areas are efficient at protecting forest areas of high and unique biodiversity in the medium term, i.e., forests will be concentrated in protected areas, where the probability  of deforestation is lower. In the long term, under a business-as-usual scenario, forests should completely disappear from protected areas while deforestation continues (Fig. \@ref(fig:fcc2100)). This phenomenon is already clearly visible in countries or states where deforestation is advanced, such as in Rondonia state (Brazil) in South America [@Ribeiro2005], Ivory Coast [@Sangne2015] or Madagascar [@Vieilledent2020] in Africa, or Cambodia [@Davis2015] in Southeast Asia. In these countries, several forested protected areas have been entirely deforested (e.g., the Haut-Sassandra protected forest in Ivory Coast, or the PK-32 Ranobe protected area in Madagascar) or severely deforested (e.g., the Beng Per wildlife sanctuary in Cambodia).

# Impact of roads on deforestation

```{r road-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="road"]
nsignif <- df$nctry_sign[df$var=="road"]
perc <- df$perc[df$var=="road"]
perc_w <- df$perc_w[df$var=="road"]
```

```{r roads-decrease, results="asis", echo=FALSE}
## Effect of roads at decreasing deforestation probability
alpha_normalized <- -2.194
coef_road_km <- -0.016 # Back-transformed parameter to have slope in km^-1
theta_mean <- inv_logit(alpha_normalized) # Mean deforestation probability
theta_road_1km <- inv_logit(alpha_normalized + coef_road_km)
d_road_1km <- 100*round(1-(theta_road_1km/theta_mean), 2)
theta_road_10km <- inv_logit(alpha_normalized + coef_road_km*10)
d_road_10km <- 100*round(1-(theta_road_10km/theta_mean), 2)
```

```{r edge-decrease, results="asis", echo=FALSE}
## Effect of edges at decreasing deforestation probability
alpha_normalized <- -2.194
coef_edge_km <- -2.736 # Back-transformed parameter to have slope in km^-1
theta_mean <- inv_logit(alpha_normalized) # Mean deforestation probability
theta_edge_1km <- inv_logit(alpha_normalized + coef_edge_km)
d_edge_1km <- 100*round(1-(theta_edge_1km/theta_mean), 2)
theta_edge_10km <- inv_logit(alpha_normalized + coef_edge_km*10)
d_edge_10km <- 100*round(1-(theta_edge_10km/theta_mean), 2)
```

Here we find that a longer distance to the road significantly reduces the risk of deforestation in `r nsignif` study areas out of `r nsa` (`r perc`% of the study areas). These `r nsignif` study areas accounted for `r perc_w`% of the tropical moist forest in 2010 (Table S7). On average, a distance of 10 km from a road reduces the risk of deforestation by `r d_road_10km`% (Fig. \@ref(fig:proba-var) and Tables S5 and S9). This said, opening a road in the forest leads to the creation of two forest edges and computing a distance from a forest pixel to the nearest road implies the existence of a distance to the forest edge. When studying the effect of roads on deforestation, it is thus impossible to neglect the effect of the distance to forest edge on the risk of deforestation.

Here, we find that the distance to the forest edge is the most important variable in determining the risk of deforestation (Table S5), in agreement with the results of other studies showing the impact of forest fragmentation on the risk of deforestation in the tropics [@Hansen2020]. We estimate that, on average, a distance of 1 km from the forest edge reduces the risk of deforestation by `r d_edge_1km`%, and a distance of 10 km reduces the risk of deforestation by almost `r d_edge_10km`% (Fig. \@ref(fig:proba-var) and Tables S5 and S9).

Consequently, building new roads in non-forest areas but close to existing forest edges would significantly increase the risk of deforestation in the nearby forest. This negative impact would be even greater if new roads are opened in the heart of forest areas. In addition to the direct deforestation associated with road building in the forest [@Kleinschroth2017], this would involve creating new forest edges and would dramatically increase deforestation probability in the area concerned. While road networks are expanding rapidly worldwide, notably in remote areas in tropical countries [@Laurance2014], our results underline the importance of conserving large roadless connected (unfragmented) forest areas.

# Uncertainty and alternative deforestation scenarios

Despite the uncertainty surrounding the mean annual deforested area for each country (Table S20), the consequences of a business-as-usual deforestation scenario on the loss of biodiversity and carbon emissions by 2100 remain clear and alarming. Moreover, given the current global context, the business-as-usual deforestation scenario we examine here appears to be very conservative. For example, we do not account for the effect of future population growth [@Raftery2012], which will likely have a major effect on deforestation, particularly in Africa, where a large part of the population depends on slash-and-burn agriculture for their livelihood [@Barnes1990; @Vieilledent2020]. Nor do we account for the increasing demand for agricultural commodities from the tropics, such as palm oil, beef and soybean, which will likely lead to a significant increase in deforestation [@Strona2018; @Karstensen2013]. Although some conservation strategies, such as protected areas, can help save some time in the fight against deforestation (being efficient at displacing deforestation toward areas of lower biodiversity or carbon stocks), it is extremely urgent to find political and socio-economic solutions that are efficient at curbing deforestation in the long term. According to previous studies [@Geist2002; @Soares-Filho2014; @Vieilledent2020; @Jones2019; @Pendrill2019], these solutions partly involve the fight against imported deforestation and an increase in international aid to tropical developing countries to help them find economic alternatives to deforestation and enforce environmental laws. We hope that the results and products of our study will facilitate the concrete application of such solutions, so that our projected deforestation map never becomes more than a scenario.
