# Introduction

Tropical forests are called the "Jewels of the Earth". They shelter 30 million species of plants and animals representing half of the Earth's wildlife and at least two-thirds of its plant species [@Gibson2011; @Wilson2012]. They also play an important role in the global carbon cycle, and at regulating the global climate [@Baccini2017; @Pan2011; @Bonan2008]. At the local scale, they regulate the regional climate [@Dickinson1992], cooling the atmosphere [@Ellison2017; @DeFrenne2019], and facilitating access to water [@Ellison2017]. They also provide protection against erosion and flooding [@Bradshaw2007]. Close to 1.6 billion people (a quarter of the world's population) rely on forest resources for their livelihoods [@FAO2015]. Despite the many ecosystem services they provide, tropical forests are disappearing at an alarming rate [@Keenan2015; @Hansen2013; @Vancutsem2020; @Achard2014], mostly because of human activites [@Curtis2018; @Geist2001]. Currently, around 8 Mha (twice the size of Switzerland) of tropical forest are disappearing each year [@FAO2015]. At this rate, will there still be tropical forests in 2100 and where will they be concentrated?

Statistical modelling of deforestation can help understand better the deforestation process, identifying the main drivers of deforestation and quantifying their relative effects. Regarding roads, modelling deforestation can help assess the impact of roads on the risk of deforestation and anticipate the changes in forest-cover associated to the potential development of road infrastructure.

Spatial forecast of tropical deforestation and inform policy makers on what's going to happen if they do nothing versus what is likely to happen in response to decisions.

In this study, we present the first attempt to model and forecast tropical deforestation globally using high resolution data.

# Spatial model of deforestation

- Some methodological information

# Effectiveness of protected areas

```{r pa-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="PA"]
nsignif <- df$nctry_sign[df$var=="PA"]
perc <- df$perc[df$var=="PA"]
perc_w <- df$perc_w[df$var=="PA"]
```

We used an extensive data-set of protected areas in the tropics (Fig. S4) obtained from the World Database on Protected Areas (<https://www.protectedplanet.net>, @WDPA2020) to test the effect of protected areas on deforestation. We considered protected areas of all IUCN categories (from Ia to VI) and of all types defined at the national level (e.g. National Parks, Reserves). 

Forested protected areas are usually placed in areas of high or unique biodiversity, in a non-random way. Protected areas are usually found in remote places with lower human disturbances and reduced accessibility, i.e. far from roads or cities, and usually at higher elevation, because low-lying arable lands have already been preempted for agriculture [@Geist2001; @Carr2004]. As a consequence, it is usually difficult to disentangle the effect of protected areas from other correlated variables, for example distance to nearest road, city, or elevation [@Andam2008]. Our multivariate logistic regression model allows disentangling the effect of each explanatory variable included in the spatial deforestation model. Moreover, the spatial random effects included in our model correct for potential bias in the protected area effect that could be associated to other unmeasured confounding variables, such as population density [@Andam2008]. 

We found that protected areas significantly reduced the risk of deforestation for `r nsignif` study-areas out of `r nsa` (`r perc`% of the study-areas). These `r nsignif` study-areas accounted for `r perc_w`% of the moist tropical forest in 2010 (Table S16). On average, protected areas reduce the risk of deforestation by 37% (Table S15). This is a clear result about the efficiency of protected areas at reducing the spatial risk of deforestation in the tropics. While other studies have assessed the effect of protected areas at reducing deforestation for particular countries [@Andam2008; @Bruner2001; @Green2013] or at efficiently protecting particular group of species [@Cazalis2020], our study is the first to assess the effect of protected areas on deforestation at the pantropical scale, using an extensive data-set on protected areas and potential confounding variables, an appropriate statistical model [@Andam2008], and accurate pantropical forest-cover change maps [@Vancutsem2020].

Our study, as other studies reporting the effect of protected areas on deforestation, demonstrates that protected areas are effective at _deplacing_ deforestation outside protected areas in tropical countries, but not necessarily that protected areas play a role at _reducing_ the deforestation intensity per se. Indeed, factors driving the intensity of deforestation at the country scale are more socio-economical or political, such as the economic development level which determines people livelihood and the link between people and deforestation [@Geist2001; @Carr2004; @Brown1994], the population size [@Barnes1990], or the environmental policy [@Soares-Filho2014]. In tropical countries with weak governance (where the environmental law enforcement is low) and with low development level (where the pressure on forest is high), it is very unlikely that protected areas will remain forested. Under a business-as-usual scenario of deforestation, we assume that the deforestation intensity will remain constant through time. When all the forest outside the protected areas will be deforested, deforestation should occur inside protected areas (Fig. \@ref(fig:fcc2100)). In this scenario, protected areas are efficient at protecting forest areas of high and unique biodiversity in the mid term. In the mid term, forest will be concentrated in protected areas, where the deforestation probability is lower. In the long term, under a business as usual scenario, forest should completely disappear from protected areas as deforestation progresses (Fig. \@ref(fig:fcc2100)). This phenomenon is already clearly observable in countries or states where deforestation is advanced, such as in the Rondonia state (Brazil) in America [@Ribeiro2005], Ivory Coast [@Sangne2015] or Madagascar [@Vieilledent2020] in Africa, or Cambodia [@Davis2015] in Asia. In these countries, several forested protected areas have been entirely (the Haut-Sassandra protected forest in Ivory-Coast, or the PK-32 Ranobe national park in Madagascar) or severely (e.g the Beng Per wildlife sanctuary in Cambodia) deforested. 

# Impact of roads on deforestation

```{r road-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="road"]
nsignif <- df$nctry_sign[df$var=="road"]
perc <- df$perc[df$var=="road"]
perc_w <- df$perc_w[df$var=="road"]
```

We also used an extensive data-set of roads in the tropics (Fig. S5) obtained from OpenStreetMap (OSM) (<https://www.openstreetmap.org/>) to estimate the effect of the distance to road on the risk of deforestation. Roads included "motorway", "trunk", "primary", "secondary" and "tertiary" roads from OSM. Distance to nearest road was computed at 150 m resolution.

We found that a higher distance to road significantly reduced the risk of deforestation for `r nsignif` study-areas out of `r nsa` (`r perc`% of the study-areas). These `r nsignif` study-areas accounted for `r perc_w`% of the moist tropical forest in 2010 (Table S16). On average, a distance of 10 km from a road reduces the risk of deforestation by 14% (Table S15). This said, when there is a road, there is a forest edge. This means that when studying the effect of roads on deforestation, it is impossible not considering also the effect of the distance to forest edge, on the deforestation risk. Indeed, opening a road in the forest implies the creation of two forest edges and computing a distance from a forest pixel to the nearest road implies the existence of a distance to forest edge. 

In our data-set, we found that the distance to forest edge was the most important variable in determining the risk of deforestation (Table S13), in line with the results of other studies showing the impact of forest fragmentation on the risk of deforestation in the tropics [@Hansen2020]. We estimated that, on average, a distance of 1 km from the forest edge reduces the risk of deforestation by 93%, and a distance of 10 km reduces the risk of deforestation by almost 100% (Table S15).

Consequently, building new roads on non-forest lands but near already existing forest edges would significantly increase the deforestation risk in the nearby forest. But this impact of roads on deforestation would be demultiplied if the development of road infrastucture implies the creation of new forest edges, in addition to the direct deforestation associated to the road building inside the forest.

Link with Laurance results

# Remaining forests in 2100

Around 7.5 Mha (75,000 km$^2$, about the size of Scotland or South Carolina) of natural old-growth moist tropical forest have been disappearing each year in the period 2010--2020.

- Countries having lost their forest
- Residual fragmented forest
- Intact forest blocks

# Impact on ecosystem services

- Impact on carbon emissions
- Biodiversity loss (map of biodiversity)
- Risk of zoonose and pandemy associated to deforestation

# Priority areas for conservation

- Refuge areas
- Impact on policies
- Various scenarios: map of probability can be used with different scenarios regarding deforestation intensity.

# Scenarios of deforestation

- Alternatives to "business-as-usual" scenarios of deforestation
- Which mean do we consider 2000--2010 or 2010--2020 or mean?
- No need to take into account diffusion of deforestation between countries and continents before 2100 (see linear change of perc with time per continent)

# Importance of assessing tropical forest resilience to climate change

[@Vieilledent2016; @Zellweger2020; @Esquivel-Muelbert2018]