# Introduction

Tropical forests are called the "Jewels of the Earth". They shelter 30 million species of plants and animals representing half of the Earth's wildlife and at least two-thirds of its plant species [@Gibson2011; @Wilson2012]. They also play an important role in the global carbon cycle, and at regulating the global climate [@Baccini2017; @Pan2011; @Bonan2008]. At the local scale, they regulate the regional climate [@Dickinson1992], cooling the atmosphere [@Ellison2017; @DeFrenne2019], and facilitating access to water [@Ellison2017]. They also provide protection against erosion and flooding [@Bradshaw2007]. Close to 1.6 billion people (a quarter of the world's population) rely on forest resources for their livelihoods [@FAO2015]. Despite the many ecosystem services they provide, tropical forests are disappearing at an alarming rate [@Keenan2015; @Hansen2013; @Vancutsem2020; @Achard2014], mostly because of human activites [@Curtis2018; @Geist2002]. Currently, around 8 Mha (twice the size of Switzerland) of tropical forest are disappearing each year [@FAO2015]. At this rate, will there still be tropical forests in 2100 and where will they be concentrated?

Statistical modelling of deforestation can help understand better the deforestation process, identifying the main drivers of deforestation and quantifying their relative effects. Regarding roads, modelling deforestation can help assess the impact of roads on the risk of deforestation and anticipate the changes in forest-cover associated to the potential development of road infrastructure.

Spatial forecast of tropical deforestation and inform policy makers on what's going to happen if they do nothing versus what is likely to happen in response to decisions.

In this study, we present the first attempt to model and forecast tropical deforestation globally using high resolution data.

# Spatial modelling of deforestation

In the present study, we modelled the spatial deforestation process for 119 study-areas representing 92 countries in the three tropical continents (America, Africa, and Asia). Study-areas covered the integrality of the tropical moist forest in the world, at the exception of some islands (eg. Sao Tome and Principe or Wallis-and-Futuna). For each study-area, we derived historical forest cover change maps on two periods of time: 1$^{\text{st}}$ January 2000 -- 1$^{\text{st}}$ January 2010, and 1$^{\text{st}}$ January 2010 -- 1$^{\text{st}}$ January 2020, from the forest cover change annual product by @Vancutsem2020 at 30 m resolution. An interactive forest cover change map for the humid tropics is available at <https://forestatrisk.cirad.fr/tropics>. For our forest definition, we only considered _natural old-growth tropical moist forests_, disregarding plantations and regrowths. We included degraded forests (not yet deforested) in our forest definition. To explain the observed deforestation on the period 2010--2020, we considered a set of spatial explanatory variables describing: topography (altitude and slope, 90 m resolution), accessibility (distances to nearest road, town, and river, 150 m resolution), forest landscape (distance to forest edge, 30 m resolution), deforestation history (distance to past deforestation, 30 m resolution), and land conservation status (presence of a protected area, 30 m resolution). This set of variables were selected on an _a priori_ knowledge of the spatial deforestation process [@Gorenflo2011; @Vieilledent2013; @Geist2002; @Brown1994; @Soares-Filho2006; @Soares-Filho2010]. Data for explanatory variables were extracted from extensive global data-sets (World Database on Protected Areas, SRTM Digital Elevation Database, and Open Street Map) and had a resolution close to the original resolution of the forest cover change map (30 m).

For each study-area, we built a large dataset from a sample of forest cover-change observations in the period 2010--2020. We performed a stratified balanced sampling between deforested and non-deforested pixels in the period 2010--2020. Pixels in each category were sampled randomly. The number of sampled observations in each study-areas was a function of the forest cover area in 2010. Datasets included between 2,701 (for Sint Maarten island in America) and 100,000 (for study-areas with high forest cover such as the Amazonas state in Brazil, Peru, DRC, and Indonesia) observations. The global data-set included a total of 3,186,698 observations: 1,601,810 of non-deforested pixels and 1,584,888 of deforested pixels, corresponding to areas of 144,163 ha and 142,647 ha, respectively.

Using sampled observations of forest cover change in the period 2010--2020, we modelled the spatial probability of deforestation as a function of the explanatory variables using a logistic regression. To account for the residual spatial variation in the deforestation process, we included additional spatial random effects for the cells of a 10 $\times$ 10 km spatial grid covering each study-area. Spatial random effects account for unmeasured or unmeasurable variables [@Clark2005] that explain a part of the residual spatial variation in the deforestation process which is not explained by the fixed spatial explanatory variables alrady included in the model (such as local population density, local environmental law enforcement, etc.). Spatial random effects were assumed spatially autocorrelated through an intrinsic conditional autoregressive (iCAR) model [@Besag1991; @Banerjee2014]. Variable selection for each study area was performed using a backward elimination procedure and parameter inference was done in a hierarchical Bayesian framework.

We compared the performance of the iCAR model at predicting the spatial probability of deforestation with three other statistical models: a null model, a simple generalized linear model (equivalent to a simple logistic regression without spatial random effects), and a Random Forest model. These two last models have been commonly used in deforestation modelling [@Soares-Filho2002; @Eastman2017; @Grinand2020; @Santos2019]. Using a cross-validation procedure, we showed that the Random Forest model was clearly overfitting the data and was less performant at predicting the probability of deforestation at new sites than the iCAR model. The iCAR model had a better predictive performance than the three other statistical models. Also, the iCAR model increased the explained deviance from 38.8 to 52.2% in average in comparison with the simple generalized linear model. This shows that environmental explanatory variables alone explain a relative small part of the spatial deforestation process, and that including spatial random effects to account for unexplained residual spatial variability strongly improves model's fit (+13.4% of deviance explained in average) and model predictive performance (+6.9% for the TSS for example). Same results were obtained when comparing accuracy indices between models at the continental scale.

Using rasters of explanatory variables at their original resolution, and the fitted iCAR model for each study-area including estimated spatial random effects, we computed the spatial probability of deforestation at 30 m resolution for the year 2020 for each study-area (Fig. \@ref(fig:prob)). An interactive global map of the spatial probability of deforestion is available at <https://forestatrisk.cirad.fr/tropics>. For each study-area, we also estimated the mean annual deforested area (in ha/yr) observed in the most recent period 2010--2020 from the forest cover change map for that period. Using the mean annual deforested area in combination with the spatial probability of deforestation map, we forecasted the forest cover change on the period 2020--2100 with a time step of 10 years, assuming a "business-as-usual" scenario. The business-as-usual scenario makes the assumption of an absence of change in both the deforestation intensity and the spatial deforestation probability in the future.

While similar studies have been performed at the scale of the Amazonian basin [@Soares-Filho2006; @Soares-Filho2010] using coarser data (1 km resolution), this is to our knowledge the first study which aims at modelling the spatial deforestation process at the pantropical scale using (i) high resolution data, (ii) extensive datasets of spatial explanatory variables, and (iii) site specific parameters. Modelling deforestation at a resolution closest to the resolution of the forest cover change map (30 m) ensures the capture of fine spatial scale deforestion processes. In particular, we demonstrated the preponderant effect of the distance to forest edge which acts at a distance much lower than 1 km on the risk of deforestation. To perform our study, we used the new `forestatrisk` Python package which has been specifically developped to model and forecast deforestation at high resolution on large spatial scales through the use of efficient numerical computation libraries such as NumPy [@Harris2020].

# Effectiveness of protected areas at deplacing deforestation

```{r pa-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="PA"]
nsignif <- df$nctry_sign[df$var=="PA"]
perc <- df$perc[df$var=="PA"]
perc_w <- df$perc_w[df$var=="PA"]
```

To test the effect of protected areas on deforestation, we used an extensive dataset of protected areas in the tropics (Fig. S4) obtained from the World Database on Protected Areas (<https://www.protectedplanet.net>, @WDPA2020). We considered protected areas of all IUCN categories (from Ia to VI) and of all types defined at the national level (e.g. National Parks, Reserves). 

Forested protected areas are usually placed in areas of high or unique biodiversity, in a non-random way. Protected areas are usually found in remote places with lower human disturbances and reduced accessibility, i.e. far from roads or cities, and usually at higher elevation, because low-lying arable lands have already been preempted for agriculture [@Geist2002; @Carr2004]. As a consequence, it is usually difficult to unravel the effect of protected areas from other correlated variables, for example distance to nearest road, city, or elevation [@Andam2008]. Our multivariate logistic regression model allows disentangling the effect of each explanatory variable included in the spatial deforestation model. Moreover, the spatial random effects included in our model correct for potential bias in the protected area effect that could be associated to other unmeasured confounding variables, such as population density [@Andam2008]. 

We found that protected areas significantly reduced the risk of deforestation for `r nsignif` study-areas out of `r nsa` (`r perc`% of the study-areas). These `r nsignif` study-areas accounted for `r perc_w`% of the tropical moist forest in 2010 (Table S16). On average, protected areas reduce the risk of deforestation by 37% (Table S15). This is a clear result demonstrating the efficiency of protected areas at reducing the spatial risk of deforestation in the tropics. While other studies have assessed the effect of protected areas at reducing deforestation for particular countries or regions [@Andam2008; @Bruner2001; @Green2013; @Soares-Filho2010] or at efficiently protecting particular group of species [@Cazalis2020], our study is the first to assess the effect of protected areas on deforestation at the pantropical scale, using an extensive dataset on protected areas and potential confounding variables, an appropriate statistical model [@Andam2008], and accurate pantropical forest-cover change maps [@Vancutsem2020].

Our study, as other studies reporting the effect of protected areas on deforestation, demonstrates that protected areas are effective at _deplacing_ deforestation outside protected areas in tropical countries, but not necessarily that protected areas play a role at _reducing_ the deforestation intensity per se. Indeed, factors driving the intensity of deforestation at the country scale are more socio-economical or political, such as the economic development level, which determines people livelihood and the link between people and deforestation [@Geist2002; @Carr2004; @Brown1994], the population size [@Barnes1990], or the environmental policy [@Soares-Filho2014]. In tropical countries with weak governance (where the environmental law enforcement is low) and with low development level (where the pressure on forest is high), it is very unlikely that protected areas will remain forested. Under a business-as-usual scenario of deforestation, we assume that the deforestation intensity will remain constant through time. When all the forest outside the protected areas will be deforested, deforestation should occur inside protected areas (Fig. \@ref(fig:fcc2100)). In this scenario, protected areas are efficient at protecting forest areas of high and unique biodiversity in the mid term. In the mid term, forest will be concentrated in protected areas, where the deforestation probability is lower. In the long term, under a business as usual scenario, forest should completely disappear from protected areas as deforestation progresses (Fig. \@ref(fig:fcc2100)). This phenomenon is already clearly observable in countries or states where deforestation is advanced, such as in the Rondonia state (Brazil) in America [@Ribeiro2005], Ivory Coast [@Sangne2015] or Madagascar [@Vieilledent2020] in Africa, or Cambodia [@Davis2015] in Asia. In these countries, several forested protected areas have been entirely (the Haut-Sassandra protected forest in Ivory-Coast, or the PK-32 Ranobe national park in Madagascar) or severely (e.g the Beng Per wildlife sanctuary in Cambodia) deforested. 

Although protected areas, being efficient at deplacing deforestation toward areas of lower biodiversity, can help us save some time in the fight against deforestation and biodiversity collapse, it is utterly urgent to find political and socio-economical solutions that are efficient at curbing the currently rampant deforestation on the long term. These solutions go in part through the fight against imported deforestation and an increase in international aid to tropical developing countries, which must be better targeted towards strong enforcement of environmental laws, education and sensibilisation of local communities to environmental stakes, the fight against corruption, the development of a more productive and environmentally friendly agriculture, and the fight against poverty [@Vieilledent2020; @Jones2019; @Soares-Filho2014; @Carr2005; @Pendrill2019; @Geist2002].

# Impact of roads on deforestation

```{r road-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="road"]
nsignif <- df$nctry_sign[df$var=="road"]
perc <- df$perc[df$var=="road"]
perc_w <- df$perc_w[df$var=="road"]
```

We also used an extensive dataset of roads in the tropics (Fig. S5) obtained from OpenStreetMap (OSM) (<https://www.openstreetmap.org/>) to estimate the effect of the distance to road on the risk of deforestation. Roads included "motorway", "trunk", "primary", "secondary", and "tertiary" roads from OSM. Distance to nearest road was computed at 150 m resolution.

We found that a higher distance to road significantly reduced the risk of deforestation for `r nsignif` study-areas out of `r nsa` (`r perc`% of the study-areas). These `r nsignif` study-areas accounted for `r perc_w`% of the tropical moist forest in 2010 (Table S16). On average, a distance of 10 km from a road reduces the risk of deforestation by 14% (Table S15). This said, where there is a road in a forest landscape, there is a forest edge. Indeed, opening a road in the forest implies the creation of two forest edges and computing a distance from a forest pixel to the nearest road implies the existence of a distance to forest edge. This means that when studying the effect of roads on deforestation, it is impossible not considering also the effect of the distance to forest edge on the deforestation risk.

In our dataset, we found that the distance to forest edge was the most important variable in determining the risk of deforestation (Table S13), in agreement with the results of other studies showing the impact of forest fragmentation on the risk of deforestation in the tropics [@Hansen2020]. We estimated that, on average, a distance of 1 km from the forest edge reduces the risk of deforestation by 93%, and a distance of 10 km reduces the risk of deforestation by almost 100% (Table S15).

Consequently, building new roads in non-forest areas but close to already existing forest edges would significantly increase the deforestation risk in the nearby forest. But this impact of roads on deforestation would be demultiplied if the development of road infrastucture is done at the heart of a forest area. This would imply the creation of new forest edges, in addition to the direct deforestation associated to the road building in the forest [@Kleinschroth2017], and would dramatically increase the deforestation probability in this area. While road networks are expanding rapidly worldwide, notably in remote areas in tropical countries [@Laurance2014], our results underline the importance of conserving large roadless and connected (unfragmented) forest areas.

# Remaining forests in 2100

```{r fcc-proj, results="asis", echo=FALSE}
## Import data
df <- read.table(here("Manuscript/Supplementary_Materials/tables/forest_cover_change.csv"), header=TRUE, sep=",")
## All study areas
nctry_loss <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::group_by(area_cont) %>%
  dplyr::summarize(nctry=n(),
                   nctry_loss21_100=sum(loss21==100),
                   nctry_loss21_75=sum(loss21>=75))
## Study-areas with more than 1 Mha of forest in 2020
nctry_loss_large <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::filter(for2020>=1e6) %>%
  dplyr::group_by(area_cont) %>%
  dplyr::summarize(nctry=n(),
                   nctry_loss21_100=sum(loss21==100),
                   nctry_loss21_75=sum(loss21>=75))
## List countries
## Study-areas with more than 1 Mha of forest in 2020
ctry <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::filter(for2020>=1e6 & loss21==100) %>%
  dplyr::arrange(area_cont, -for2020)
```

Around 7.5 Mha (75,000 km$^2$, about the size of Scotland or South Carolina) of natural old-growth tropical moist forest have been disappearing each year in the 2010--2020 last decade. At this rate, 53% of the world tropical moist forest will have disappeared over the 21$^{st}$ century (Table \@ref(tab:fcc-reg)). This average at the global scale is hidding large differences in the deforestation rates at the continental (Table \@ref(tab:fcc-reg) and Fig. S11) and country (Table S8 and S10) scales. In Asia, where the remaining forest in 2020 is relatively low (256 Mha) and the annual deforested areas are relatively large (2.5 Mha/yr) compared with the two other continents, the percentage of forest cover loss over the 21$^{st}$ century should reach 75%. In Africa, where the annual deforested areas are lower (2.1 Mha/yr), this percentage should be of 58%. In America, where the annual deforested areas are higher (2.9 Mha/yr), but where the remaining forest in 2020 is much higher than in Asia and Africa (634 Mha), this percentage should be of 41%. Under a business-as-usual scenario of deforestation, i.e. in the absence of any change that could efficiently curb deforestation in the future, most (75%) of the tropical moist forests existing in 2000 will have disappeared around 2100, 2150, and 2200 in Asia, Africa, and America, respectively (Table \@ref(tab:fcc-reg) and Fig. S11).

At the country scale, 59 study areas (33 in America, 19 in Africa, and 7 in Asia) out of 119 will experience a complete loss of their tropical forests before 2100 (Table S10 and Fig. \@ref(fig:fcc2100)). When considering only study-areas with more than 1 Mha of forest in 2020, this represents 21 study-areas (9 in America, 7 in Africa, and 5 in Asia) out of 63. These 21 study-areas include most of the countries of Central America (Mexico, Guatemala, Honduras, Nicaragua), of Western (Sierra Leone, Ivory Coast, Ghana, Nigeria) and Eastern (Ethiopia, Madagascar) Africa, and of continental Asia (Myanmar, Laos, Vietnam, Cambodia, Indian Western Ghats). These countries represent almost all of 5 biodiversity hotspots identified by @Myers2000: Mesoamerica, West African forests, Madagascar, South Central China, Western Ghats and Sri Lanka.

Regarding the remaining tropical moist forest in 2100, three large blocks of relatively intact forest should emerge (Fig. \@ref(fig:fcc2100) and interactive map of projected deforestation in 2020--2100 available at <https://forestatrisk.cirad.fr/tropics>). A first forest block will be located in America and will include the upper part of the Amazonian Basin (including the Peruvian, Ecuadorian, Colombian and Venezuelan Amazonia) and the Guyana Plateau (Guyana, Suriname, and French Guiana). A second one will be located in West Central Africa and will include forests from Gabon, Equatorial Guinea, Cameroon, Central African Republic, and Congo. A third one will be located in Melanesia and will include forests from Papua New Guinea, Solomon Islands and Vanuatu.

Apart from these three large and relatively intact forest blocks, the remaining tropical moist forest in 2100 will be highly fragmented (Fig. \@ref(fig:fcc2100)). In America, highly fragmented forest will be found in the Brazilian states of the Amazonian deforestation arc (Acre, Rondonia, Mato Grosso, Para, Amapa) and in the Roraima state at the North of Amazonia. In Africa, forests of the Democratic Republic of the Congo will also be highly fragmented and will be completely separated from the large forest block at the West of the Congo River. In Asia, small patches of heavily fragmented forests will remain in Malaysia, Indonesia and in the Philippines (Fig. \@ref(fig:fcc2100)). Remaining forests will be concentrated in remote areas (far from roads and villages), preferentially in protected areas, and at high elevation. For example, remaining forests in Borneao will be concentrated in the Betung Kerihun and Kayan Mantarang National Parks.

As tropical forests shelter most of the terrestrial biodiversity and stock most of the carbon on land, projected deforestation will likely have dramatical effects on future biodiversity loss and climate change. While the impact of projected deforestation on carbon emissions will be discussed below, completing a rigourous assessment of the impact of the projected deforestation on biodiversity is out of the scope of this study. Such analysis would require accurate species distribution and biodiversity maps including a high number of species representative of the biodiversity in the tropics. Such maps are not available to date. But if we only consider the species 

# Carbon emissions under a business-as-usual scenario of deforestation

We estimated the carbon emissions associated to projected deforestation in the period 2020--2100 under a "business-as-usual" scenario of deforestation. We used a time step of 10 years to project the change in carbon emissions with time. To do so, we used the pantropical 1 km resolution aboveground dry biomass (AGB, in Mg/ha) map by @Avitabile2016 (Fig. S12). Biomass can be converted into carbon emissions using the Intergovernmental Panel on Climate Change (IPCC) default carbon fraction of 0.47 [@McGroddy2004]. Although a large uncertainty surrounds the biomass estimates, Avitabile's map is the most accurate biomass map available at the pantropical scale, and represents correctly the aboveground biomass spatial variability. We assumed no change of the forest carbon stocks in the future while computing carbon emissions associated to projected deforestation. Under a "business-as-usual" scenario of deforestation (no change in the annual deforested area, in ha/yr, in the future), the change in mean annual carbon emissions in the future is only attributable to the spatial variation of the forest carbon stocks and to the location of the future deforestation.

We found that carbon emissions associated to deforestation of tropical moist forests at the global scale should increase from 0.54 Pg/yr in 2010--2020 up to 0.73 Pg/yr in 2080--2090 (+0.19 Pg/yr corresponding to a 35% increase, Fig \@ref(fig:c-em)). This increase in carbon emissions was predicted for the three continents (Fig \@ref(fig:c-em)). A decrease of the carbon emissions was predicted for Asia from the period 2070--2080 and at the global scale from the period 2090--2100 (Fig \@ref(fig:c-em)).

Increase in carbon emissions is explained by the fact that forest areas which are progressively deforested have higher carbon stocks. Several previous studies have shown that elevation is an important variable determining forest carbon stocks [@Vieilledent2016; @Keith2009; @Stegen2011; @Saatchi2011; @Baccini2012]. Forest carbon stocks are expected to be optimal at mid elevation [@Vieilledent2016; @Asner2012], because of higher orographic precipitation, and because the climatic stress associated to strong wind and low temperature is lower at mid elevation than at high elevation. In the present study, we show that low-elevation areas are preferentially deforested in comparison with high-elevation areas (Tables S12 and S13). This is explained by the fact that low-elevation areas are more accessible and because more arable lands are concentrated at low elevation, where the slope is usually lower and the soil is more productive [@Geist2002; @Carr2004]. Consequently, increase in carbon emissions in the future can be explained by the fact that deforestation progresses towards higher elevation areas where forest carbon stocks are higher. Moreover, remote forest areas, that have been less disturbed by human activities in the past [@Potapov2017] can have accumulated large quantities of carbon [@Dargie2017; @Brinck2017]. The progressive deforestation of more intact forests can also explain the projected increase in carbon emissions in the future.

The decrease in carbon emissions predicted for Asia from the period 2070--2080, and at the global scale from the period 2090--2100, could be associated to either a decrease of the carbon stocks of deforested areas (in association with environmental factors, such as lower carbon stocks at high elevation for example), or to a decrease of the total deforested area at the continental and global scale as countries progressively loss all their forests. Vietnam and Laos should experience a complete loss of their tropical moist forests during the decade 2070--2080, and for Myanmar, this should happen during the decade 2080--2090 (Table S10). Because these three countries account for a significant part (22%) of the annual deforested area in Asia (550,568 ha/yr out of 2,472,128 ha/yr, see Table S8 and S9), carbon emissions are expected to decrease for Asia from 2070. Nevertheless, no other country should loose its forest in Asia during the decade 2090--2100 (apart from Singapore which has a very small annual deforested areas of 172 ha/yr, see Table S8) so the strong decrease of carbon emissions in Asia for that period is only attributable to the carbon stock spatial variation.

Our estimates of 0.54 Pg/yr (0.22, 0.21, and 0.12 Pg/yr for America, Asia, and Africa, respectively) of carbon emissions associated to tropical deforestation for the period 2010--2020 are consistent with previous studies [@Baccini2017; @Harris2012; @Achard2014]. For the period 2000-2014, previous studies found estimates of 0.81-0.88 Pg/yr of carbon emissions associated to deforestation when considering all tropical forest types (both moist and dry tropical forests), while our study focuses on tropical moist forests only. The projected 35% increase in carbon emissions associated to tropical deforestation should strongly modify the role of tropical forests in the global carbon cycle. Under a business-as-usual scenario of deforestation, tropical forests will likely act as an increasing carbon source [@Baccini2017], thus reinforcing climate change in the future.

These results demonstrate the importance of having spatial predictions of the deforestation to forecast the carbon emissions associated to deforestation. Under a business-as-usual scenario of deforestation, carbon emissions are the result of both the spatial variation of the forest carbon stocks and the location of the future deforestation. The projected increase in carbon emission estimates with time would have been hard to anticipate with mean annual forest cover change estimates by country and forest type (such as those provided by the Forest Resource Assessment report for example, @Keenan2015) and mean forest carbon stocks by continent and forest type (such as the emission factors provided by @IPCC2019).

# Alternative deforestation scenarios and research pespectives

## Scenarios of deforestation

- Alternatives to "business-as-usual" scenarios of deforestation
- Which mean do we consider 2000--2010 or 2010--2020 or mean?
- No need to take into account diffusion of deforestation between countries and continents before 2100 (see linear change of perc with time per continent)

## Impact on ecosystem services

- Impact on carbon emissions
- Biodiversity loss (map of biodiversity)
- Risk of zoonose and pandemy associated to deforestation

## Priority areas for conservation

- Refuge areas
- Impact on policies
- Various scenarios: map of probability can be used with different scenarios regarding deforestation intensity.

## Importance of assessing tropical forest resilience to climate change

[@Vieilledent2016; @Zellweger2020; @Esquivel-Muelbert2018]