# Introduction

Tropical forests are at the heart of concerns when it comes to climate change and biodiversity loss, which both pose unprecedented threats to human civilization [@IPCC2014; @IPBES2019; @Cardinale2012]. Commonly called the "Jewels of the Earth", tropical forests shelter 30 million species of plants and animals representing half of the Earth's wildlife and at least two-thirds of its plant species [@Gibson2011; @Wilson2012]. Through carbon sequestration, they play an important role in the global carbon cycle, and in regulating the global climate [@Baccini2017; @Pan2011; @Bonan2008]. Intact tropical forest ecosystems also prevent zoonoses outbreak and decrease the risk of pandemics [@IPBES2020; @Dobson2020; @Tollefson2020]. At the local scale, tropical forests regulate the regional climate [@Dickinson1992], cooling the atmosphere [@Ellison2017; @DeFrenne2019], and facilitate access to water [@Ellison2017]. They also provide protection against erosion and flooding [@Bradshaw2007]. Close to 1.6 billion people (a quarter of the world's population) rely on forest resources for their livelihoods [@FAO2015]. Despite the many ecosystem services they provide, tropical forests are disappearing at an alarming rate [@Keenan2015; @Hansen2013; @Vancutsem2020; @Achard2014], mostly because of human activities [@Curtis2018; @Geist2002]. Past studies have estimated that about 8 Mha of tropical forest (including both moist and dry forests) are currently disappearing each year [@FAO2015]. At this rate, will there still be tropical forests at the end of the 21^st^ century, where will they be concentrated, and what will be the consequences of tropical deforestation on climate and biodiversity in the future?

Forecasting forest cover change is paramount as it allows to anticipate the consequences of deforestation (in terms of carbon emissions, biodiversity loss, or water supply) under various technological, political and socio-economic scenarios, and inform decision makers accordingly [@Clark2001]. The "business-as-usual" scenario (i.e. in the absence of any change in deforestation in the future) is of particular interest as it allows to anticipate the likely future in the absence of change, and thus alert decision-makers, if necessary, to an essential change of course in order to avoid any potential environmental disaster. If scenarios of carbon dioxide emissions and climate change have been developed for several years by the Intergovernmental Panel on Climate Change [@IPCC2014], and are now largely used by the scientific community and known to the general public, there is currently no equivalent scenarios at the global scale for land-use change and biodiversity [@IPBES2019]. Spatialized scenarios of forest cover change are particularly important because both forest carbon stocks [@Avitabile2016; @Baccini2017] and biodiversity [@Allnutt2008; @Kremen2008; @Mittermeier2011] vary considerably in space at fine scale. Non-spatial scenarios of forest cover change [@FAO2015] cannot be used to accurately forecast associated carbon emissions and change in biodiversity, nor for systematic conservation planning at the local scale. Spatial forecasts of forest cover change rest on spatial statistical models which allow estimating a probability of change in space as a function of a set of spatial predictors. In addition to the forecasts, models can be used to identify the main drivers of deforestation and quantify their relative effects. For example, models can be used to assess the impact of roads on the risk of deforestation [@Laurance2014] and the effectiveness of protected areas at reducing deforestation [@Andam2008].

\newpage

Few studies have attempted to provide spatialized forest cover change scenarios in the tropics on large spatial scales. Largest studies to date have focused on modelling and forecasting forest cover change at the scale of the Amazonian basin [@Soares-Filho2006; @Soares-Filho2010]. In this study, we present the first attempt to model and forecast tropical deforestation at the global scale using high resolution data. This was made possible by the recent availability of pantropical forest/tree cover change datasets at high resolution [@Hansen2013; @Vancutsem2020] and the existence of high-resolution global datasets of spatial factors of deforestation (World Database on Protected Areas, SRTM Digital Elevation Database, and OpenStreetMap). Using these extensive datasets, we used an advanced statistical approach to model the deforestation spatially in the humid tropics. We tested the effectiveness of protected areas at reducing deforestation and assess the impact of roads on forest conservation globally. We derived high resolution global maps of the deforestation risk and likely future forest cover for the 21^st^ century under a business-as-usual scenario, including an uncertainty analysis. We also estimated the carbon emissions associated to projected deforestation.

# Remaining forests in 2100

```{r fcc-proj, results="asis", echo=FALSE}
require(dplyr)
require(here)
## Import data
df <- read.table(here::here("Manuscript/Supplementary_Materials/tables/forest_cover_change_mean.csv"), header=TRUE, sep=",")
## All study areas
nctry_loss <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::group_by(area_cont) %>%
  dplyr::summarize(nctry=n(),
                   nctry_loss21_100=sum(loss21==100),
                   nctry_loss21_75=sum(loss21>=75))
## Study-areas with more than 1 Mha of forest in 2020
nctry_loss_large <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::filter(for2020>=1e6) %>%
  dplyr::group_by(area_cont) %>%
  dplyr::summarize(nctry=n(),
                   nctry_loss21_100=sum(loss21==100),
                   nctry_loss21_75=sum(loss21>=75))
## List countries
## Study-areas with more than 1 Mha of forest in 2020
ctry <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::filter(for2020>=1e6 & loss21==100) %>%
  dplyr::arrange(area_cont, -for2020)
```

We show here that around 7.6 Mha (76,000 km^2^, about the size of Scotland or South Carolina) of natural old-growth tropical moist forest have been disappearing each year during the last decade (2010-2020). At this rate, 61% of the world's tropical moist forest will have disappeared over the 21^st^ century (Table \@ref(tab:fcc-reg)). This average at the global scale is hiding large differences in the deforestation rates at the continental (Table \@ref(tab:fcc-reg) and Fig. S12) and country (Tables S14 and S16) scales. In Asia, where the remaining forest area in 2020 is relatively low (247 Mha) and the annual deforested area is relatively large (2.5 Mha/yr) compared with the two other continents, the percentage of forest cover loss over the 21^st^ century should reach 78%. In Africa, where the annual deforested area is lower (2.2 Mha/yr), this percentage should be 58%. In America, where the annual deforested area is higher (2.9 Mha/yr), but where the remaining forest in 2020 is much higher than in Asia and Africa (621 Mha), this percentage should be 41%. Under a business-as-usual scenario of deforestation, most (75%) of the tropical moist forests existing in 2000 will have disappeared around 2100, 2150, and 2200 in Asia, Africa, and America, respectively (Table \@ref(tab:fcc-reg) and Fig. S12).

At the country scale, 47 countries (18 in America, 21 in Africa, and 8 in Asia) out of the 92 we studied, plus 18 states and 1 region in Brazil and India, respectively, will experience a complete loss of their tropical forests before 2100 (Table S16 and Fig. \@ref(fig:fcc2100)). When considering only countries with more than 1 Mha of forest in 2020, this represents 20 countries (6 in America, 8 in Africa, and 6 in Asia) out of 48, plus 5 states and 1 region in Brazil and India, respectively. These 20 countries include most of the countries of Central America (Mexico, Guatemala, Honduras, Nicaragua), of Western (Sierra Leone, Ivory Coast, Ghana, Nigeria) and Eastern (Ethiopia, Madagascar) Africa, and of continental Asia (Myanmar, Laos, Vietnam, Cambodia, Malaysia, Indian Western Ghats). These regions include almost all the tropical forest of 6 biodiversity hotspots identified by @Mittermeier2011: Mesoamerica, Guinean Forests of West Africa, Horn of Africa, Madagascar and Indian Ocean Islands, Indo-Burma, Western Ghats and Sri Lanka.

Using an advanced spatial statistical modelling approach (see Methods), we obtained high resolution (30 m) pantropical maps of the deforestation risk (Fig. \@ref(fig:prob)), and of the projected forest cover for the 21^st^ century in the humid tropics (Fig. \@ref(fig:fcc2100)), under a business-as-usual scenario. Regarding the remaining tropical moist forest in 2100, three large blocks of relatively intact forest will emerge (see both Fig. \@ref(fig:fcc2100) and the interactive map of projected deforestation for the period 2020--2100 available at <https://forestatrisk.cirad.fr/maps.html>). A first forest block will be located in America and will include the upper part of the Amazonian Basin (including the Peruvian, Ecuadorian, Colombian and Venezuelan Amazonia) and the Guyana Plateau (Guyana, Suriname, and French Guiana). A second one will be located in West Central Africa and will include forests from Gabon, Equatorial Guinea, Cameroon, Central African Republic, and Congo. A third one will be located in Melanesia and will include forests from Papua New Guinea, Solomon Islands, and Vanuatu.

Apart from these three large and relatively intact forest blocks, the remaining tropical moist forest in 2100 will be highly fragmented (Fig. \@ref(fig:fcc2100)). In America, highly fragmented forest will be found in the Brazilian states of the Amazonian deforestation arc (Acre, Rondonia, Mato Grosso, Para, Amapa) and in the Roraima state in the North of Amazonia. In Africa, forests of the Democratic Republic of the Congo will also be highly fragmented and will be completely separated from the large forest block located at the West of the Congo River. In Asia, small patches of heavily fragmented forests will remain in Thailand, Indonesia and in the Philippines (Fig. \@ref(fig:fcc2100)). Remaining forests will be concentrated in remote areas (far from roads and villages), preferentially in protected areas, and at high elevation. For example, remaining forests in Borneo will be concentrated in the Betung Kerihun and Kayan Mentarang National Parks.

As tropical forests shelter most of the terrestrial biodiversity and stock most of the carbon on land, projected deforestation will likely have dramatical effects on future biodiversity loss and climate change. While the impact of projected deforestation on carbon emissions will be discussed below, completing a rigorous assessment of the impact of the projected deforestation on biodiversity is out of the scope of this study. Such analysis would require accurate species distribution and biodiversity maps including a high number of species representative of the biodiversity in the tropics [@Allnutt2008]. Such maps are not available to date (but see @Jetz2012, @Jenkins2013, or @Pimm2014). Nonetheless, as a first estimate, if we consider only the endemic species [@Mittermeier2011] from the 6 biodiversity hotspots where almost all the tropical forest is predicted to disappear by 2100, and assuming that most of these species depend on tropical moist forests, deforestation would induce the extinction of 29,140 species of plants and 4,576 species of vertebrates (including birds, reptiles, amphibians, freshwater fishes, and mammals) which cannot be found anywhere else on Earth (Table S18).

# Carbon emissions under a business-as-usual scenario of deforestation

We estimated the carbon emissions associated with projected deforestation in the period 2020--2100 under a business-as-usual scenario of deforestation. We assumed no change of the forest carbon stocks in the future when computing carbon emissions associated with projected deforestation. Under a business-as-usual scenario of deforestation (no change in the annual deforested area, in ha/yr, in the future), the change in mean annual carbon emissions in the future is only attributable to the spatial variation of the forest carbon stocks and to the location of the future deforestation. We found that carbon emissions associated to deforestation of tropical moist forests at the global scale should increase from 0.525 Pg/yr in 2010--2020 up to 0.746 Pg/yr in 2080--2090 (+0.221 Pg/yr corresponding to a 42% increase, Fig \@ref(fig:c-em)). This increase in carbon emissions was predicted for the three continents (Fig \@ref(fig:c-em)). A decrease of the carbon emissions was predicted from the period 2070--2080 for Asia and from the period 2080--2090 at the global scale (Fig \@ref(fig:c-em)).

Increase in carbon emissions is explained by the fact that forest areas which are progressively deforested have higher carbon stocks. Several previous studies have shown that elevation is an important variable determining forest carbon stocks [@Vieilledent2016; @Keith2009; @Stegen2011; @Saatchi2011; @Baccini2012]. Forest carbon stocks are expected to be optimal at mid elevation [@Vieilledent2016; @Asner2012], because of higher orographic precipitation, and because the climatic stress associated with strong wind and low temperature is lower at mid elevation than at high elevation. In the present study, we show that low-elevation areas are preferentially deforested in comparison to high-elevation areas (Tables S4 and S5). This is because low-elevation areas are more accessible to human populations and because more arable lands are concentrated at low elevation, where the slope is usually lower and the soil is more productive [@Geist2002; @Carr2004]. Consequently, increase in carbon emissions in the future can be explained by the fact that deforestation progresses towards higher elevation areas where forest carbon stocks are higher. Moreover, remote forest areas that have been less disturbed by human activities in the past [@Potapov2017] can have accumulated large quantities of carbon [@Dargie2017; @Brinck2017]. The progressive deforestation of more intact forests can also explain the projected increase in carbon emissions in the future.

The decrease in carbon emissions predicted from the period 2070--2080 for Asia, and from the period 2080--2090 at the global scale, could potentially be associated to either a decrease of the carbon stocks of deforested areas (in association with environmental factors, such as lower carbon stocks at high elevation for example), or to a decrease of the total deforested area at the continental and global scale, as countries progressively lose all their forest. In Asia, six countries will experience a complete loss of their forest between 2070 and 2100 (Table S16). These countries (which include Laos, Malaysia, Myanmar, and Vietnam), account for a significant part (32%) of the annual deforested area in Asia (808,375 ha/yr out of 2,532,927 ha/yr, see Tables S14 and S15). This explains for a great part the predicted decrease of carbon emissions in Asia from 2070.

Our estimates of 0.525 Pg/yr (0.208, 0.199, and 0.118 Pg/yr for America, Asia, and Africa, respectively) of carbon emissions associated to tropical deforestation for the period 2010--2020 are consistent with previous studies [@Baccini2017; @Harris2012; @Achard2014]. For the period 2000-2014, previous studies found estimates of 0.81-0.88 Pg/yr of carbon emissions associated with deforestation when considering all tropical forest types (both moist and dry tropical forests), while our study focuses on tropical _moist_ forests only. Because of the projected 42% increase in carbon emissions associated with tropical deforestation under a business-as-usual scenario, tropical forests will act as an increasing net carbon source [@Baccini2017], thus reinforcing climate change in the future.

These results demonstrate the importance of having spatial predictions of deforestation to forecast carbon emissions associated with deforestation. Under a business-as-usual scenario of deforestation, carbon emissions are the result of both the spatial variation of the forest carbon stocks and the location of the future deforestation. The projected increase in carbon emission estimates with time would have been hard to anticipate with mean annual forest cover change estimates by country and forest type (such as those provided by the Forest Resource Assessment report for example, @Keenan2015) and mean forest carbon stocks by continent and forest type (such as the emission factors provided by @IPCC2019).

# Effectiveness of protected areas at displacing deforestation

```{r pa-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="PA"]
nsignif <- df$nctry_sign[df$var=="PA"]
perc <- df$perc[df$var=="PA"]
perc_w <- df$perc_w[df$var=="PA"]
```

```{r pa-decrease, results="asis", echo=FALSE}
## Inv_logit function
inv_logit <- function (x, min=0, max=1) {
    p <- exp(x)/(1 + exp(x))
    p <- ifelse(is.na(p) & !is.na(x), 1, p)
    p * (max - min) + min
}
## Effect of protected areas at decreasing deforestation probability
alpha_normalized <- -2.194
coef_pa_km <- -0.483 # Back-transformed parameter to have slope in km^-1
theta_mean <- inv_logit(alpha_normalized) # Mean deforestation probability
theta_pa <- inv_logit(alpha_normalized + coef_pa_km)
d_pa <- 100*round(1-(theta_pa/theta_mean), 2)
```

Forested protected areas are usually placed in areas of high or unique biodiversity, in a non-random way. They are also usually found in remote places with lower human disturbances and reduced accessibility, i.e. far from roads or cities (Fig. S6), and usually at higher elevation, because low-lying arable lands have already been preempted for agriculture [@Geist2002; @Carr2004]. As a consequence, it is usually difficult to unravel the effect of protected areas from other correlated variables, for example distance to nearest road, city, or elevation [@Andam2008]. The multivariate logistic regression model used in our study allows disentangling the effect of each explanatory variable included in the spatial deforestation model. Moreover, the spatial random effects included in our model are correcting for potential bias in the protected area effect that could be associated with other unmeasured confounding variables, such as population density [@Andam2008].

We found that protected areas significantly reduced the risk of deforestation for `r nsignif` study-areas out of `r nsa` (`r perc`% of the study-areas). These `r nsignif` study-areas accounted for `r perc_w`% of the tropical moist forest in 2010 (Table S6). On average, protected areas reduce the risk of deforestation by `r d_pa`% (Fig. \@ref(fig:proba-var) and Tables S5 and S9). This is a clear result demonstrating the efficiency of protected areas at reducing the spatial risk of deforestation in the tropics. While other studies have assessed the effect of protected areas at reducing deforestation for particular countries or regions [@Andam2008; @Bruner2001; @Green2013; @Soares-Filho2010] or at efficiently protecting particular group of species [@Cazalis2020], our study is the first to assess the effect of protected areas on deforestation at the pantropical scale, using an extensive dataset on protected areas and potential confounding variables, an appropriate statistical model [@Andam2008], and accurate pantropical forest-cover change maps [@Vancutsem2020].

Our study, as other studies reporting the effect of protected areas on deforestation, demonstrates that protected areas are effective at _displacing_ deforestation outside protected areas in tropical countries, but not necessarily that protected areas play a role at _reducing_ the deforestation intensity per se. Indeed, factors driving the intensity of deforestation at the country scale are more socio-economic or political factors, such as the economic development level, which determines people livelihood and the link between people and deforestation [@Geist2002; @Carr2004; @Brown1994], the population size [@Barnes1990], or the environmental policy [@Soares-Filho2014]. In tropical countries with weak governance (where the environmental law enforcement is low) and with a low development level (where the pressure on forest is high), it is very unlikely that protected areas will remain forested. Under a business-as-usual scenario of deforestation, we assume that the deforestation intensity will remain constant through time. When all the forest outside the protected areas will be deforested, deforestation should occur inside protected areas (Fig. \@ref(fig:fcc2100)). In this scenario, protected areas are efficient at protecting forest areas of high and unique biodiversity in the mid term. In the mid term, forest will be concentrated in protected areas, where the deforestation probability is lower. In the long term, under a business as usual scenario, forest should completely disappear from protected areas as deforestation continues (Fig. \@ref(fig:fcc2100)). This phenomenon is already clearly observable in countries or states where deforestation is advanced, such as in the Rondonia state (Brazil) in America [@Ribeiro2005], Ivory Coast [@Sangne2015] or Madagascar [@Vieilledent2020] in Africa, or Cambodia [@Davis2015] in Asia. In these countries, several forested protected areas have been entirely (the Haut-Sassandra protected forest in Ivory-Coast, or the PK-32 Ranobe national park in Madagascar) or severely (e.g the Beng Per wildlife sanctuary in Cambodia) deforested.

# Impact of roads on deforestation

```{r road-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="road"]
nsignif <- df$nctry_sign[df$var=="road"]
perc <- df$perc[df$var=="road"]
perc_w <- df$perc_w[df$var=="road"]
```

```{r roads-decrease, results="asis", echo=FALSE}
## Effect of roads at decreasing deforestation probability
alpha_normalized <- -2.194
coef_road_km <- -0.016 # Back-transformed parameter to have slope in km^-1
theta_mean <- inv_logit(alpha_normalized) # Mean deforestation probability
theta_road_1km <- inv_logit(alpha_normalized + coef_road_km)
d_road_1km <- 100*round(1-(theta_road_1km/theta_mean), 2)
theta_road_10km <- inv_logit(alpha_normalized + coef_road_km*10)
d_road_10km <- 100*round(1-(theta_road_10km/theta_mean), 2)
```

```{r edge-decrease, results="asis", echo=FALSE}
## Effect of edges at decreasing deforestation probability
alpha_normalized <- -2.194
coef_edge_km <- -2.736 # Back-transformed parameter to have slope in km^-1
theta_mean <- inv_logit(alpha_normalized) # Mean deforestation probability
theta_edge_1km <- inv_logit(alpha_normalized + coef_edge_km)
d_edge_1km <- 100*round(1-(theta_edge_1km/theta_mean), 2)
theta_edge_10km <- inv_logit(alpha_normalized + coef_edge_km*10)
d_edge_10km <- 100*round(1-(theta_edge_10km/theta_mean), 2)
```

We found that a higher distance to the road significantly reduced the risk of deforestation for `r nsignif` study-areas out of `r nsa` (`r perc`% of the study-areas). These `r nsignif` study-areas accounted for `r perc_w`% of the tropical moist forest in 2010 (Table S7). On average, a distance of 10 km from a road reduces the risk of deforestation by `r d_road_10km`% (Fig. \@ref(fig:proba-var) and Tables S5 and S9). This said, where there is a road in a forest landscape, there is a forest edge. Indeed, opening a road in the forest implies the creation of two forest edges and computing a distance from a forest pixel to the nearest road implies the existence of a distance to forest edge. This means that when studying the effect of roads on deforestation, it is impossible not considering also the effect of the distance to forest edge on the deforestation risk.

In our dataset, we found that the distance to forest edge was the most important variable in determining the risk of deforestation (Table S5), in agreement with the results of other studies showing the impact of forest fragmentation on the risk of deforestation in the tropics [@Hansen2020]. We estimated that, on average, a distance of 1 km from the forest edge reduces the risk of deforestation by `r d_edge_1km`%, and a distance of 10 km reduces the risk of deforestation by almost `r d_edge_10km`% (Fig. \@ref(fig:proba-var) and Tables S5 and S9).

Consequently, building new roads in non-forest areas but close to already existing forest edges would significantly increase the deforestation risk in the nearby forest. But this negative impact would be substantially stronger if the development of roads is done at the heart of forest areas. In addition to the direct deforestation associated with road building in the forest [@Kleinschroth2017], this would imply the creation of new forest edges, and would dramatically increase the deforestation probability in this area. While road networks are expanding rapidly worldwide, notably in remote areas in tropical countries [@Laurance2014], our results underline the importance of conserving large roadless and connected (unfragmented) forest areas.

# Uncertainty and alternative deforestation scenarios

Despite an uncertainty surrounding the mean annual deforested area for each country (Table S20), the consequences of a business-as-usual scenario of deforestation on the loss of biodiversity and carbon emissions by 2100 remain clear and alarming. Moreover, given the current global context, the business-as-usual scenario of deforestation we examined appears to be very conservative. For example, we did not take into account the effect of population growth in the future [@Raftery2012], which will likely have a strong effect on deforestation, particularly in Africa, where a large part of the population depends on slash-and-burn agriculture for their livelihood [@Barnes1990; @Vieilledent2020]. Nor we accounted for the increasing demand for agricultural commodities from the tropics, such as oil palm, beef meat or soybean, which will likely lead to a significant increase in deforestation [@Strona2018; @Karstensen2013]. Although some conservation strategies, such as protected areas, can help us save some time in the fight against deforestation (being efficient at displacing deforestation toward areas of lower biodiversity or carbon stocks), it is utterly urgent to find political and socio-economical solutions that are efficient at curbing deforestation on the long term. According to previous studies [@Vieilledent2020; @Jones2019; @Soares-Filho2014; @Carr2005; @Pendrill2019; @Geist2002], these solutions partly involve the fight against imported deforestation, and an increase in international aid to tropical developing countries to assist them find economic alternatives to deforestation and enforce the environmental law. We hope that the results and products of our study will help to implement such solutions in a concrete way, so that our projected deforestation map remains as a scenario.
