# Introduction

Climate change and biodiversity loss pose unprecedented threats to human civilization [@IPCC2014; @IPBES2019; @Cardinale2012]. Tropical forests are at the heart of concerns when it comes to climate change and biodiversity loss. Commonly called the "Jewels of the Earth", they shelter 30 million species of plants and animals representing half of the Earth's wildlife and at least two-thirds of its plant species [@Gibson2011; @Wilson2012]. Through carbon sequestration, they play an important role in the global carbon cycle, and in regulating the global climate [@Baccini2017; @Pan2011; @Bonan2008]. Intact tropical forest ecosystems also prevent zoonoses outbreak and decrease the risk of pandemics [@IPBES2020; @Dobson2020; @Tollefson2020]. At the local scale, tropical forests regulate the regional climate [@Dickinson1992], cooling the atmosphere [@Ellison2017; @DeFrenne2019], and facilitate access to water [@Ellison2017]. They also provide protection against erosion and flooding [@Bradshaw2007]. Close to 1.6 billion people (a quarter of the world's population) rely on forest resources for their livelihoods [@FAO2015]. Despite the many ecosystem services they provide, tropical forests are disappearing at an alarming rate [@Keenan2015; @Hansen2013; @Vancutsem2020; @Achard2014], mostly because of human activities [@Curtis2018; @Geist2002]. Currently, around 8 Mha (twice the size of Switzerland) of tropical forest are disappearing each year [@FAO2015]. At this rate, will there still be tropical forests in 2100, where will they be concentrated, and what will be the consequences of tropical deforestation on climate and biodiversity in the future?

Forecasting forest cover change is paramount as it allows to anticipate the consequences of deforestation (in terms of carbon emissions, biodiversity loss, or water supply) under various technological, political and socio-economic scenarios, and inform decision makers accordingly [@Clark2001]. The business-as-usual scenario is of particular interest as it allows to anticipate the likely future in the absence of change, and thus alert decision-makers, if necessary, to an essential change of course in order to avoid any potential environmental disaster. If scenarios of carbon dioxide emissions and climate change have been developed for several years by the Intergovernmental Panel on Climate Change [@IPCC2014], and are now largely used by the scientific community and known to the general public, there is currently no equivalent scenarios at the global scale for land-use change and biodiversity [@IPBES2019]. Spatialized scenarios of forest cover change are particularly important because both forest carbon stocks [@Avitabile2016; @Baccini2017] and biodiversity [@Allnutt2008; @Kremen2008; @Mittermeier2011] vary considerably in space at fine scale. Non-spatial scenarios of forest cover change [@FAO2015] cannot be used to accurately forecast associated carbon emissions and change in biodiversity, nor for systematic conservation planning at the local scale. Spatial forecasts of forest cover change rest on spatial statistical models which allow estimating a probability of change in space as a function of a set of spatial predictors. In addition to the forecasts, models can be used to identify the main drivers of deforestation and quantify their relative effects. For example, models can be used to assess the impact of roads on the risk of deforestation [@Laurance2014] and the effectiveness of protected areas at reducing deforestation [@Andam2008].

Few studies have attempted to provide spatialized forest cover change scenarios in the tropics on large spatial scales. Largest studies to date have focused on modelling forest cover change at the scale of the Amazonian basin [@Soares-Filho2006; @Soares-Filho2010]. In this study, we present the first attempt to model and forecast tropical deforestation at the global scale using high resolution data. This is made possible by the recent availability of pantropical forest or tree cover change datasets at high resolution [@Hansen2013; @Vancutsem2020] and the existence of high-resolution global datasets on spatial factors of deforestation (World Database on Protected Areas, SRTM Digital Elevation Database, and OpenStreetMap). Using these extensive datasets, we modelled the deforestation spatially in the humid tropics. We demonstrated the effectiveness of protected areas at displacing deforestation and the negative impact of roads on forest conservation globally. We derived high resolution global maps of the deforestation risk and likely future forest cover for the 21$^{\text{st}}$ century under a "business-as-usual" scenario. If large areas of tropical forest will still remain in 2100 at the heart of the Amazon, in the Congo Basin, and in New-Guinea, several countries will experience a complete loss of their tropical forests by 2100. Remaining forests will be strongly fragmented and located in remote places. As deforestation should progress toward areas of higher carbon stocks, carbon emissions associated to deforestation of tropical moist forests at the global scale should increase from 0.54 Pg/yr in 2010--2020 up to 0.73 Pg/yr in 2080--2090, thus reinforcing the carbon source effect of tropical forests in the future [@Baccini2017].

# Spatial modelling of deforestation

In the present study, we modelled the spatial deforestation process for 119 study-areas representing 92 countries in the three tropical continents (America, Africa, and Asia). Study-areas covered the integrality of the tropical moist forest in the world, at the exception of some islands (eg. Sao Tome and Principe or Wallis-and-Futuna). For each study-area, we derived historical forest cover change maps on two periods of time: 1$^{\text{st}}$ January 2000 -- 1$^{\text{st}}$ January 2010, and 1$^{\text{st}}$ January 2010 -- 1$^{\text{st}}$ January 2020, from the forest cover change annual product by @Vancutsem2020 at 30 m resolution. An interactive forest cover change map for the humid tropics is available at <https://forestatrisk.cirad.fr/tropics>. For our forest definition, we only considered _natural old-growth tropical moist forests_, disregarding plantations and regrowths. We included degraded forests (not yet deforested) in our forest definition. To explain the observed deforestation on the period 2010--2020, we considered a set of spatial explanatory variables describing: topography (altitude and slope, 90 m resolution), accessibility (distances to nearest road, town, and river, 150 m resolution), forest landscape (distance to forest edge, 30 m resolution), deforestation history (distance to past deforestation, 30 m resolution), and land conservation status (presence of a protected area, 30 m resolution). This set of variables were selected on an _a priori_ knowledge of the spatial deforestation process [@Gorenflo2011; @Vieilledent2013; @Geist2002; @Brown1994; @Soares-Filho2006; @Soares-Filho2010]. Data for explanatory variables were extracted from extensive global data-sets (World Database on Protected Areas, SRTM Digital Elevation Database, and OpenStreetMap) and had a resolution close to the original resolution of the forest cover change map (30 m).

For each study-area, we built a large dataset from a sample of forest cover change observations in the period 2010--2020. We performed a stratified balanced sampling between deforested and non-deforested pixels in the period 2010--2020. Pixels in each category were sampled randomly. The number of sampled observations in each study-areas was a function of the forest cover area in 2010. Datasets included between 2,701 (for Sint Maarten island in America) and 100,000 (for study-areas with high forest cover such as the Amazonas state in Brazil, Peru, DRC, and Indonesia) observations. The global data-set included a total of 3,186,698 observations: 1,601,810 of non-deforested pixels and 1,584,888 of deforested pixels, corresponding to areas of 144,163 ha and 142,647 ha, respectively.

Using sampled observations of forest cover change in the period 2010--2020, we modelled the spatial probability of deforestation as a function of the explanatory variables using a logistic regression. To account for the residual spatial variation in the deforestation process, we included additional spatial random effects for the cells of a 10 $\times$ 10 km spatial grid covering each study-area. Spatial random effects account for unmeasured or unmeasurable variables [@Clark2005] that explain a part of the residual spatial variation in the deforestation process which is not explained by the fixed spatial explanatory variables already included in the model (such as local population density, local environmental law enforcement, etc.). Spatial random effects were assumed spatially autocorrelated through an intrinsic conditional autoregressive (iCAR) model [@Besag1991; @Banerjee2014]. Variable selection for each study area was performed using a backward elimination procedure and parameter inference was done in a hierarchical Bayesian framework.

We compared the performance of the iCAR model at predicting the spatial probability of deforestation with three other statistical models: a null model, a simple generalized linear model (equivalent to a simple logistic regression without spatial random effects), and a Random Forest model. These two last models have been commonly used for deforestation modelling [@Soares-Filho2002; @Eastman2017; @Grinand2020; @Santos2019]. Using a cross-validation procedure, we showed that the Random Forest model was clearly overfitting the data and was less performant at predicting the probability of deforestation at new sites than the iCAR model. The iCAR model had a better predictive performance than the three other statistical models. Also, the iCAR model increased the explained deviance from 38.8 to 52.2% in average in comparison with the simple generalized linear model. This shows that environmental explanatory variables alone explain a relative small part of the spatial deforestation process, and that including spatial random effects to account for unexplained residual spatial variability strongly improves model's fit (+13.4% of deviance explained in average) and model predictive performance (+6.9% for the TSS for example). Same results were obtained when comparing accuracy indices between models at the continental scale.

Using rasters of explanatory variables at their original resolution, and the fitted iCAR model for each study-area including estimated spatial random effects, we computed the spatial probability of deforestation at 30 m resolution for the year 2020 for each study-area (Fig. \@ref(fig:prob)). An interactive global map of the spatial probability of deforestation is available at <https://forestatrisk.cirad.fr/tropics>. For each study-area, we also estimated the mean annual deforested area (in ha/yr) observed in the most recent period 2010--2020 from the forest cover change map for that period. Using the mean annual deforested area in combination with the spatial probability of deforestation map, we forecasted the forest cover change on the period 2020--2100 with a time step of 10 years, assuming a "business-as-usual" scenario. The business-as-usual scenario makes the assumption of an absence of change in both the deforestation intensity and the spatial deforestation probability in the future.

While similar studies have been performed at the scale of the Amazonian basin [@Soares-Filho2006; @Soares-Filho2010] using coarser data (1 km resolution), this is to our knowledge the first study which aims at modelling the spatial deforestation process at the pantropical scale using (i) high resolution data, (ii) extensive datasets of spatial explanatory variables, and (iii) site specific parameters. Modelling deforestation at a resolution closest to the resolution of the forest cover change map (30 m) ensures the capture of fine spatial scale deforestation processes. In particular, we demonstrated the preponderant effect of the distance to forest edge which acts at a distance much lower than 1 km on the risk of deforestation (Fig. \@ref(fig:proba-var)). To perform our study, we used the new `forestatrisk` Python package which has been specifically developed to model and forecast deforestation at high resolution on large spatial scales through the use of efficient numerical computation libraries such as NumPy [@Harris2020].

# Effectiveness of protected areas at displacing deforestation

```{r pa-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="PA"]
nsignif <- df$nctry_sign[df$var=="PA"]
perc <- df$perc[df$var=="PA"]
perc_w <- df$perc_w[df$var=="PA"]
```

To test the effect of protected areas on deforestation, we used an extensive dataset of protected areas in the tropics (Fig. S4) obtained from the World Database on Protected Areas (<https://www.protectedplanet.net>, @WDPA2020). We considered protected areas of all IUCN categories (from Ia to VI) and of all types defined at the national level (e.g. National Parks, Reserves).

Forested protected areas are usually placed in areas of high or unique biodiversity, in a non-random way. Protected areas are usually found in remote places with lower human disturbances and reduced accessibility, i.e. far from roads or cities (Fig. S4), and usually at higher elevation, because low-lying arable lands have already been preempted for agriculture [@Geist2002; @Carr2004]. As a consequence, it is usually difficult to unravel the effect of protected areas from other correlated variables, for example distance to nearest road, city, or elevation [@Andam2008]. Our multivariate logistic regression model allows disentangling the effect of each explanatory variable included in the spatial deforestation model. Moreover, the spatial random effects included in our model are correcting for potential bias in the protected area effect that could be associated with other unmeasured confounding variables, such as population density [@Andam2008].

We found that protected areas significantly reduced the risk of deforestation for `r nsignif` study-areas out of `r nsa` (`r perc`% of the study-areas). These `r nsignif` study-areas accounted for `r perc_w`% of the tropical moist forest in 2010 (Table S16). On average, protected areas reduce the risk of deforestation by 37% (Fig. \@ref(fig:proba-var) and Table S15). This is a clear result demonstrating the efficiency of protected areas at reducing the spatial risk of deforestation in the tropics. While other studies have assessed the effect of protected areas at reducing deforestation for particular countries or regions [@Andam2008; @Bruner2001; @Green2013; @Soares-Filho2010] or at efficiently protecting particular group of species [@Cazalis2020], our study is the first to assess the effect of protected areas on deforestation at the pantropical scale, using an extensive dataset on protected areas and potential confounding variables, an appropriate statistical model [@Andam2008], and accurate pantropical forest-cover change maps [@Vancutsem2020].

Our study, as other studies reporting the effect of protected areas on deforestation, demonstrates that protected areas are effective at _displacing_ deforestation outside protected areas in tropical countries, but not necessarily that protected areas play a role at _reducing_ the deforestation intensity per se. Indeed, factors driving the intensity of deforestation at the country scale are more socio-economic or political, such as the economic development level, which determines people livelihood and the link between people and deforestation [@Geist2002; @Carr2004; @Brown1994], the population size [@Barnes1990], or the environmental policy [@Soares-Filho2014]. In tropical countries with weak governance (where the environmental law enforcement is low) and with a low development level (where the pressure on forest is high), it is very unlikely that protected areas will remain forested. Under a business-as-usual scenario of deforestation, we assume that the deforestation intensity will remain constant through time. When all the forest outside the protected areas will be deforested, deforestation should occur inside protected areas (Fig. \@ref(fig:fcc2100)). In this scenario, protected areas are efficient at protecting forest areas of high and unique biodiversity in the mid term. In the mid term, forest will be concentrated in protected areas, where the deforestation probability is lower. In the long term, under a business as usual scenario, forest should completely disappear from protected areas as deforestation progresses (Fig. \@ref(fig:fcc2100)). This phenomenon is already clearly observable in countries or states where deforestation is advanced, such as in the Rondonia state (Brazil) in America [@Ribeiro2005], Ivory Coast [@Sangne2015] or Madagascar [@Vieilledent2020] in Africa, or Cambodia [@Davis2015] in Asia. In these countries, several forested protected areas have been entirely (the Haut-Sassandra protected forest in Ivory-Coast, or the PK-32 Ranobe national park in Madagascar) or severely (e.g the Beng Per wildlife sanctuary in Cambodia) deforested.

Although protected areas, being efficient at displacing deforestation toward areas of lower biodiversity, can help us save some time in the fight against deforestation and biodiversity collapse, it is utterly urgent to find political and socio-economical solutions that are efficient at curbing the currently rampant deforestation on the long term. These solutions go in part through the fight against imported deforestation and an increase in international aid to tropical developing countries, which must be better targeted towards strong enforcement of environmental laws, education and sensibilisation of local communities to environmental stakes, the fight against corruption, the development of a more productive and environmentally friendly agriculture, and the fight against poverty [@Vieilledent2020; @Jones2019; @Soares-Filho2014; @Carr2005; @Pendrill2019; @Geist2002].

# Impact of roads on deforestation

```{r road-sign, results="asis", echo=FALSE}
## Import data
f <- here("Manuscript", "Supplementary_Materials", "tables", "sign_PA_road.csv")
df <- read.table(f, header=TRUE, sep=",")
nsa <- df$nctry[df$var=="road"]
nsignif <- df$nctry_sign[df$var=="road"]
perc <- df$perc[df$var=="road"]
perc_w <- df$perc_w[df$var=="road"]
```

We also used an extensive dataset of roads in the tropics (Fig. S5) obtained from OpenStreetMap (OSM) (<https://www.openstreetmap.org/>) to estimate the effect of the distance to road on the risk of deforestation. Roads included "motorway", "trunk", "primary", "secondary", and "tertiary" roads from OSM. Distance to the nearest road was computed at 150 m resolution.

We found that a higher distance to the road significantly reduced the risk of deforestation for `r nsignif` study-areas out of `r nsa` (`r perc`% of the study-areas). These `r nsignif` study-areas accounted for `r perc_w`% of the tropical moist forest in 2010 (Table S16). On average, a distance of 10 km from a road reduces the risk of deforestation by 14% (Fig. \@ref(fig:proba-var) and Table S15). This said, where there is a road in a forest landscape, there is a forest edge. Indeed, opening a road in the forest implies the creation of two forest edges and computing a distance from a forest pixel to the nearest road implies the existence of a distance to forest edge. This means that when studying the effect of roads on deforestation, it is impossible not considering also the effect of the distance to forest edge on the deforestation risk.

In our dataset, we found that the distance to forest edge was the most important variable in determining the risk of deforestation (Table S13), in agreement with the results of other studies showing the impact of forest fragmentation on the risk of deforestation in the tropics [@Hansen2020]. We estimated that, on average, a distance of 1 km from the forest edge reduces the risk of deforestation by 93%, and a distance of 10 km reduces the risk of deforestation by almost 100% (Fig. \@ref(fig:proba-var) and Table S15).

Consequently, building new roads in non-forest areas but close to already existing forest edges would significantly increase the deforestation risk in the nearby forest. But this negative impact is substantially stronger if the development of roads is done at the heart of a forest area. In addition to the direct deforestation associated with road building in the forest [@Kleinschroth2017], this would imply the creation of new forest edges, and would dramatically increase the deforestation probability in this area. While road networks are expanding rapidly worldwide, notably in remote areas in tropical countries [@Laurance2014], our results underline the importance of conserving large roadless and connected (unfragmented) forest areas.

# Remaining forests in 2100

```{r fcc-proj, results="asis", echo=FALSE}
## Import data
df <- read.table(here("Manuscript/Supplementary_Materials/tables/forest_cover_change_mean.csv"), header=TRUE, sep=",")
## All study areas
nctry_loss <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::group_by(area_cont) %>%
  dplyr::summarize(nctry=n(),
                   nctry_loss21_100=sum(loss21==100),
                   nctry_loss21_75=sum(loss21>=75))
## Study-areas with more than 1 Mha of forest in 2020
nctry_loss_large <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::filter(for2020>=1e6) %>%
  dplyr::group_by(area_cont) %>%
  dplyr::summarize(nctry=n(),
                   nctry_loss21_100=sum(loss21==100),
                   nctry_loss21_75=sum(loss21>=75))
## List countries
## Study-areas with more than 1 Mha of forest in 2020
ctry <- df %>%
  dplyr::mutate(loss21=round(100*(for2000-for2100)/for2000)) %>%
  dplyr::filter(for2020>=1e6 & loss21==100) %>%
  dplyr::arrange(area_cont, -for2020)
```

Around 7.5 Mha (75,000 km$^2$, about the size of Scotland or South Carolina) of natural old-growth tropical moist forest have been disappearing each year in the 2010--2020 last decade. At this rate, 53% of the world's tropical moist forest will have disappeared over the 21$^{st}$ century (Table \@ref(tab:fcc-reg)). This average at the global scale is hiding large differences in the deforestation rates at the continental (Table \@ref(tab:fcc-reg) and Fig. S11) and country (Table S8 and S10) scales. In Asia, where the remaining forest in 2020 is relatively low (256 Mha) and the annual deforested areas are relatively large (2.5 Mha/yr) compared with the two other continents, the percentage of forest cover loss over the 21$^{st}$ century should reach 75%. In Africa, where the annual deforested areas are lower (2.1 Mha/yr), this percentage should be 58%. In America, where the annual deforested areas are higher (2.9 Mha/yr), but where the remaining forest in 2020 is much higher than in Asia and Africa (634 Mha), this percentage should be 41%. Under a business-as-usual scenario of deforestation, i.e. in the absence of any change that could efficiently curb deforestation in the future, most (75%) of the tropical moist forests existing in 2000 will have disappeared around 2100, 2150, and 2200 in Asia, Africa, and America, respectively (Table \@ref(tab:fcc-reg) and Fig. S11).

At the country scale, 59 study areas (33 in America, 19 in Africa, and 7 in Asia) out of 119 will experience a complete loss of their tropical forests before 2100 (Table S10 and Fig. \@ref(fig:fcc2100)). When considering only study-areas with more than 1 Mha of forest in 2020, this represents 21 study-areas (9 in America, 7 in Africa, and 5 in Asia) out of 63. These 21 study-areas include most of the countries of Central America (Mexico, Guatemala, Honduras, Nicaragua), of Western (Sierra Leone, Ivory Coast, Ghana, Nigeria) and Eastern (Ethiopia, Madagascar) Africa, and of continental Asia (Myanmar, Laos, Vietnam, Cambodia, Indian Western Ghats). These countries represent almost all the tropical forest of 6 biodiversity hotspots identified by @Mittermeier2011: Mesoamerica, Guinean Forests of West Africa, Horn of Africa, Madagascar and Indian Ocean Islands, Indo-Burma, Western Ghats and Sri Lanka.

Regarding the remaining tropical moist forest in 2100, three large blocks of relatively intact forest should emerge (Fig. \@ref(fig:fcc2100) and interactive map of projected deforestation in 2020--2100 available at <https://forestatrisk.cirad.fr/tropics>). A first forest block will be located in America and will include the upper part of the Amazonian Basin (including the Peruvian, Ecuadorian, Colombian and Venezuelan Amazonia) and the Guyana Plateau (Guyana, Suriname, and French Guiana). A second one will be located in West Central Africa and will include forests from Gabon, Equatorial Guinea, Cameroon, Central African Republic, and Congo. A third one will be located in Melanesia and will include forests from Papua New Guinea, Solomon Islands and Vanuatu.

Apart from these three large and relatively intact forest blocks, the remaining tropical moist forest in 2100 will be highly fragmented (Fig. \@ref(fig:fcc2100)). In America, highly fragmented forest will be found in the Brazilian states of the Amazonian deforestation arc (Acre, Rondonia, Mato Grosso, Para, Amapa) and in the Roraima state in the North of Amazonia. In Africa, forests of the Democratic Republic of the Congo will also be highly fragmented and will be completely separated from the large forest block at the West of the Congo River. In Asia, small patches of heavily fragmented forests will remain in Malaysia, Indonesia and in the Philippines (Fig. \@ref(fig:fcc2100)). Remaining forests will be concentrated in remote areas (far from roads and villages), preferentially in protected areas, and at high elevation. For example, remaining forests in Borneo will be concentrated in the Betung Kerihun and Kayan Mentarang National Parks.

As tropical forests shelter most of the terrestrial biodiversity and stock most of the carbon on land, projected deforestation will likely have dramatical effects on future biodiversity loss and climate change. While the impact of projected deforestation on carbon emissions will be discussed below, completing a rigorous assessment of the impact of the projected deforestation on biodiversity is out of the scope of this study. Such analysis would require accurate species distribution and biodiversity maps including a high number of species representative of the biodiversity in the tropics [@Allnutt2008]. Such maps are not available to date (but see @Jetz2012, @Jenkins2013, or @Pimm2014). Nonetheless, as a first estimate, if we only consider the endemic species [@Mittermeier2011] from the 6 biodiversity hotspots where almost all the tropical forest is predicted to disappear by 2100, and assuming that most of these species depend on tropical moist forests, deforestation would induce the extinction of 29,140 species of plants and 4,576 species of vertebrates (including birds, reptiles, amphibians, freshwater fishes, and mammals) which cannot be found anywhere else on Earth.

# Carbon emissions under a business-as-usual scenario of deforestation

We estimated the carbon emissions associated with projected deforestation in the period 2020--2100 under a "business-as-usual" scenario of deforestation. We used a time step of 10 years to project the change in carbon emissions with time. To do so, we used the pantropical 1 km resolution aboveground dry biomass (AGB, in Mg/ha) map by @Avitabile2016 (Fig. S12). Biomass can be converted into carbon emissions using the Intergovernmental Panel on Climate Change (IPCC) default carbon fraction of 0.47 [@McGroddy2004]. Although a large uncertainty surrounds the biomass estimates, Avitabile's map is the most accurate biomass map available at the pantropical scale, and represents correctly the aboveground biomass spatial variability. We assumed no change of the forest carbon stocks in the future while computing carbon emissions associated with projected deforestation. Under a "business-as-usual" scenario of deforestation (no change in the annual deforested area, in ha/yr, in the future), the change in mean annual carbon emissions in the future is only attributable to the spatial variation of the forest carbon stocks and to the location of the future deforestation.

We found that carbon emissions associated to deforestation of tropical moist forests at the global scale should increase from 0.54 Pg/yr in 2010--2020 up to 0.73 Pg/yr in 2080--2090 (+0.19 Pg/yr corresponding to a 35% increase, Fig \@ref(fig:c-em)). This increase in carbon emissions was predicted for the three continents (Fig \@ref(fig:c-em)). A decrease of the carbon emissions was predicted for Asia from the period 2070--2080 and at the global scale from the period 2090--2100 (Fig \@ref(fig:c-em)).

Increase in carbon emissions is explained by the fact that forest areas which are progressively deforested have higher carbon stocks. Several previous studies have shown that elevation is an important variable determining forest carbon stocks [@Vieilledent2016; @Keith2009; @Stegen2011; @Saatchi2011; @Baccini2012]. Forest carbon stocks are expected to be optimal at mid elevation [@Vieilledent2016; @Asner2012], because of higher orographic precipitation, and because the climatic stress associated with strong wind and low temperature is lower at mid elevation than at high elevation. In the present study, we show that low-elevation areas are preferentially deforested in comparison with high-elevation areas (Tables S12 and S13). This is explained by the fact that low-elevation areas are more accessible and because more arable lands are concentrated at low elevation, where the slope is usually lower and the soil is more productive [@Geist2002; @Carr2004]. Consequently, increase in carbon emissions in the future can be explained by the fact that deforestation progresses towards higher elevation areas where forest carbon stocks are higher. Moreover, remote forest areas that have been less disturbed by human activities in the past [@Potapov2017] can have accumulated large quantities of carbon [@Dargie2017; @Brinck2017]. The progressive deforestation of more intact forests can also explain the projected increase in carbon emissions in the future.

The decrease in carbon emissions predicted for Asia from the period 2070--2080, and at the global scale from the period 2090--2100, could be associated to either a decrease of the carbon stocks of deforested areas (in association with environmental factors, such as lower carbon stocks at high elevation for example), or to a decrease of the total deforested area at the continental and global scale as countries progressively loss all their forests. Vietnam and Laos should experience a complete loss of their tropical moist forests during the decade 2070--2080, and for Myanmar, this should happen during the decade 2080--2090 (Table S10). Because these three countries account for a significant part (22%) of the annual deforested area in Asia (550,568 ha/yr out of 2,472,128 ha/yr, see Tables S8 and S9), carbon emissions are expected to decrease for Asia from 2070. Nevertheless, no other country should lose its forest in Asia during the decade 2090--2100 (apart from Singapore which has a very small annual deforested areas of 172 ha/yr, see Table S8) so the strong decrease of carbon emissions in Asia for that period is only attributable to the carbon stock spatial variation.

Our estimates of 0.54 Pg/yr (0.22, 0.21, and 0.12 Pg/yr for America, Asia, and Africa, respectively) of carbon emissions associated to tropical deforestation for the period 2010--2020 are consistent with previous studies [@Baccini2017; @Harris2012; @Achard2014]. For the period 2000-2014, previous studies found estimates of 0.81-0.88 Pg/yr of carbon emissions associated with deforestation when considering all tropical forest types (both moist and dry tropical forests), while our study focuses on tropical moist forests only. The projected 35% increase in carbon emissions associated with tropical deforestation should strongly modify the role of tropical forests in the global carbon cycle. Under a business-as-usual scenario of deforestation, tropical forests will likely act as an increasing carbon source [@Baccini2017], thus reinforcing climate change in the future.

These results demonstrate the importance of having spatial predictions of deforestation to forecast the carbon emissions associated with deforestation. Under a business-as-usual scenario of deforestation, carbon emissions are the result of both the spatial variation of the forest carbon stocks and the location of the future deforestation. The projected increase in carbon emission estimates with time would have been hard to anticipate with mean annual forest cover change estimates by country and forest type (such as those provided by the Forest Resource Assessment report for example, @Keenan2015) and mean forest carbon stocks by continent and forest type (such as the emission factors provided by @IPCC2019).

# Alternative deforestation scenarios and research perspectives

Our global map of the deforestation risk can be used to predict future forest cover change under alternative scenarios of deforestation intensity. In our case, we considered a business-as-usual scenario using the mean annual deforested area for each study-area in the last decade 2010--2020 (7.5 Mha/yr at the global scale). This appears to be a very conservative scenario. First, the deforestation intensity can increase depending on the reference period chosen for the scenario. Considering the period 2000--2020, the mean annual deforested area would be 8.3 Mha/yr at the global scale (Table S9). Second, under the business-as-usual scenario, we do not take into account the effect of population growth which could potentially have a strong effect on deforestation in Africa [@Barnes1990; @Raftery2012], or the increasing demand for agricultural commodities from the tropics, such as oil palm, beef meat or soybean, which could potentially have a strong impact on deforestation throughout the tropics [@Strona2018; @Karstensen2013]. Our predictions show that under a business-as-usual scenario, large areas of forests will remain in 2100. This underlines the importance of studying the resilience of tropical forests to climate change by 2100 as remaining forests could be potentially severely threatened by climate change [@Vieilledent2016; @Esquivel-Muelbert2018]. The high-resolution pantropical maps we produced can be used both at the global scale for developing new biodiversity and carbon emission scenarios [@IPCC2014; @IPBES2019] and at the local scale for implementing new conservation strategies including the REDD+ (Reducing Emissions from Deforestation and Forest Degradation) mechanism or systematic conservation planning through the identification of areas at high risk of deforestation and refuge areas for biodiversity in the future.
